---
title: "FSLANAT and FSLVBM pipeline"
author: "by Elise Delzant - `r format(Sys.time(), '%d %B %Y')`"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  epuRate::PCTG:
    toc: TRUE
    number_sections: FALSE
    code_folding: "show"
---


```{r, echo=FALSE}
options(width = 60)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

```

# Create bulk file

```{bash, message=FALSE, eval=FALSE}

bind="/network/lustre/iss02/ukbiobank/software" 
wd="/network/lustre/iss02/ukbiobank/unorganized/unorganized_pheno"
cd $wd
${bind}/ukbconv ukb50815.enc_ukb bulk -s20253


```

# Extract first visits (remove imaging second wave)

```{R, message=FALSE, eval=FALSE}

blk=read.table("ukb50815_20252.bulk", stringsAsFactors = F)

table(blk$V2)
write.table(blk[-which(blk$V2=="20252_3_0"),], "ukb50815.20252_2.bulk", col.names = F, row.names = F, quote=F)

```

#Remove unusable subjects (that I gotten through the FSL process (_20252))

```{R, message =FALSE, eval=FALSE}
blk_v2=read.table("ukb50815.20252_2.bulk",stringsAsFactors = F)
unusable=read.table("unusable.csv",header=F)
usables=setdiff(ukb$V1,unusable$V1)
ukb_usable=ukb[ukb$V1 %in%usables,]
write.table(ukb_usable,'ukb50815.20252.bulk',col.names=F,row.names=F,quote=F)

```

#before
```{bash, message=FALSE, eval=FALSE}

bind="/network/lustre/iss02/ukbiobank/software" 
ukb="/network/lustre/iss02/ukbiobank/unorganized/unorganized_pheno" # UKB files
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"

mkdir -p ${wd}/Downloadfsl
cp ${ukb}/k53185r50815.key ${wd}/Downloadfsl/k53185r50815.key
cp ${wd}/ukb50815.20252.bulk  ${wd}/Downloadfsl/ukb50815.20252.bulk

bind="/network/lustre/iss02/ukbiobank/software" 
ukb="/network/lustre/iss02/ukbiobank/unorganized/unorganized_pheno" # UKB files
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"



cd /network/lustre/iss02/home/elise.delzant/Downloadfsl
bind="/network/lustre/iss02/ukbiobank/software" 

for step in $(seq 1 1 1)   # ....
do 
${bind}/ukbfetch -bdiff.bulk -ak53185r673035.key -s${step} -m1 #new key 
done 
```


#step 0 create directory 
```{bash, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
wd='/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl'
ukb="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
cd ${ukb}
mkdir -p FSLVBM_files
mkdir -p FSLVBM


cd ${ukb}/FSLVBM_files

mkdir -p Template_anat
mkdir -p Template_vbm

cd ${ukb}/FSLVBM_files/Template_vbm
echo -n >mergelist_vbm 
echo -n >mergelist_vbm2


cd ${ukb}/FSLVBM_files/Template_anat
echo -n >mergelist_anat
echo -n >mergelist_anat2

ukb="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
cd ${ukb}/Downloadfsl
echo *.zip> id.csv
```

#convert array to table
```{r, echo=FALSE,message=FALSE,warning=FALSE,eval=F}
id<-read.table("id.csv",sep=" ",header=FALSE)
write.table(t(id),"../../../id.csv",row.names = FALSE,col.names = FALSE)
getwd()
```


#step 1-2ab : brain extraction and tissue seg for vbm, and Affine-registration to the GM ICBM-152 template, concatenation and average for both 

```{bash, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
#!/bin/bash
#SBATCH --partition=bigmem
#SBATCH --job-name=fsl
#SBATCH --time=00:25:00
#SBATCH --mem=1G
#SBATCH --chdir=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl
#SBATCH --ntasks=1
#SBATCH --array=1-1
#SBATCH --output=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/fsl.log
#SBATCH --error=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/fsl.err
#SBATCH --cpus-per-task=1

module load FSL

T=${FSLDIR}data/standard/tissuepriors/avg152T1_gray
wd='/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfs'
ukb="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
fsl=${ukb}/FSLVBM
fsl_f=${ukb}/FSLVBM_files
cd ${wd}
export participants=id.csv
dat_0=$(sed -n ${SLURM_ARRAY_TASK_ID}p $participants)

dat=`echo ${dat_0}| xargs`

subject=$(echo $dat | cut -d'_' -f 1)
echo ${subject}

pdir=${wd}/${subject}
mkdir -p ${pdir} 
unzip ${wd}/${dat} -d ${pdir}

fslsub=${fsl}/${subject}
mkdir -p ${fslsub}'/anat'
mkdir -p ${fslsub}'/vbm'

mv ${pdir}'/T1/T1_fast/T1_brain_pve_1.nii.gz' ${fslsub}'/anat'
mv ${pdir}'/T1/T1_fast/T1_brain_pve_2.nii.gz' ${fslsub}'/anat'
mv ${pdir}'/T1/T1_brain.nii.gz' ${fslsub}'/vbm'

#for the fslanat pipeline we rename the image and register image to grey matter template T
cd ${fslsub}'/anat'
for g in `${FSLDIR}bin/imglob *brain_pve_1*` 
do
${FSLDIR}bin/immv ${g} ${g}_struc
done

for g in `${FSLDIR}bin/imglob *_struc.*`
do
${FSLDIR}bin/fsl_reg ${g} $T ${g}_GM_to_T -a
done

name=`echo *_struc.* | sed 's/_struc././g'` 
root=`$FSLDIR/bin/remove_ext $name | sort -u`
path=${fslsub}/anat/${root}_struc_GM_to_T
cd ${fsl_f}/Template_anat
echo ${path} >>mergelist_anat

#for the fslvbm pipeline fslvbm1_bet , creates a 'struc' file with renamed images of brain extraction
cd ${fslsub}'/vbm'
fslvbm_1_bet -b

#step 2a for this pipeline, brain extraction and tissue segmentation 
cd struc
for g in `$FSLDIR/bin/imglob *_struc.*` 
do
$FSLDIR/bin/fast -R 0.3 -H 0.1 ${g}_brain 
$FSLDIR/bin/immv ${g}_brain_pve_1 ${g}_GM 
done

for g in `${FSLDIR}bin/imglob *_struc.*`
do
${FSLDIR}bin/fsl_reg ${g}_GM $T ${g}_GM_to_T -a
done

name=`echo *_struc.* | sed 's/_struc././g'` 
root=`$FSLDIR/bin/remove_ext $name | sort -u`
path=${fslsub}/vbm/struc/${root}_struc_GM_to_T
cd ${fsl_f}/Template_vbm
echo ${path} >>mergelist_vbm

cd ${wd}
rm -r ${subject}
rm ${subject}'_20252_2_0.zip'
```

#step 2c : Averaged image is flipped along x-axis and the two mirror images then re-averaged  first-pass template 
```{bash, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
#!/bin/bash
#SBATCH --partition=bigmem
#SBATCH --job-name=fsl
#SBATCH --time=02:00:00
#SBATCH --mem=4G
#SBATCH --chdir=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl
#SBATCH --ntasks=1
#SBATCH --output=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/template1.log
#SBATCH --error=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/template1.err
#SBATCH --cpus-per-task=1

tps=$(date)
echo $tps
module load FSL
wd='/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl'
ukb="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
fsl=${ukb}/FSLVBM
fsl_f=${ukb}/FSLVBM_files
cd ${fsl_f}/Template_anat

mergelist=`cat mergelist_anat`

${FSLDIR}bin/fslmerge -t template_4D_GM $mergelist
${FSLDIR}bin/fslmaths template_4D_GM -Tmean template_GM
${FSLDIR}bin/fslswapdim template_GM -x y z template_GM_flipped
${FSLDIR}bin/fslmaths template_GM -add template_GM_flipped -div 2 template_GM_init

cd ${fsl_f}/Template_vbm
mergelist=`cat mergelist_vbm`
${FSLDIR}bin/fslmerge -t template_4D_GM $mergelist
${FSLDIR}bin/fslmaths template_4D_GM -Tmean template_GM
${FSLDIR}bin/fslswapdim template_GM -x y z template_GM_flipped
${FSLDIR}bin/fslmaths template_GM -add template_GM_flipped -div 2 template_GM_init

tps_fin=$(date)
echo $tps_fin
```

#step 2d: Non-linear Registration to first pass template ,concatenation and average
```{bash, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
#!/bin/bash
#SBATCH --partition=bigmem
#SBATCH --job-name=fsl
#SBATCH --time=00:20:00
#SBATCH --mem=1G
#SBATCH --chdir=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl
#SBATCH --ntasks=1
#SBATCH --array=1-400
#SBATCH --output=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/fsl.log
#SBATCH --error=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/fsl.err
#SBATCH --cpus-per-task=1

module load FSL 

wd='/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl'
ukb="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
fsl=${ukb}/FSLVBM
fsl_f=${ukb}/FSLVBM_files
T_anat=${fsl_f}/Template_anat/template_GM_init 
T_vbm=${fsl_f}/Template_vbm/template_GM_init

cd ${wd}
export participants=id.csv

dat_0=$(sed -n ${SLURM_ARRAY_TASK_ID}p $participants)
dat=`echo ${dat_0}| xargs`
subject=$(echo $dat | cut -d'_' -f 1)

fslsub=${fsl}/${subject}

#fslvbm2d for pipeline anat 
cd ${fslsub}'/anat'
for g in `$FSLDIR/bin/imglob *_struc.*`
do 
${FSLDIR}/bin/fsl_reg ${g} $T_anat ${g}_GM_to_T_init -fnirt "--config=GM_2_MNI152GM_2mm.cnf"
done 


name=`echo *_struc.* | sed 's/_struc././g'` 
root=`$FSLDIR/bin/remove_ext $name | sort -u`
path=${fslsub}/anat/${root}_struc_GM_to_T_init
cd ${fsl_f}/Template_anat
echo ${path} >>mergelist_anat2

#fslvbm2d for pipeline vbm

cd ${fslsub}'/vbm'
cd struc 
for g in `$FSLDIR/bin/imglob *_struc.*`
do 
${FSLDIR}/bin/fsl_reg ${g}_GM $T_vbm ${g}_GM_to_T_init -fnirt "--config=GM_2_MNI152GM_2mm.cnf"
done 

name=`echo *_struc.* | sed 's/_struc././g'` 
root=`$FSLDIR/bin/remove_ext $name | sort -u`
path=${fslsub}/vbm/struc/${root}_struc_GM_to_T_init
cd ${fsl_f}/Template_vbm
echo ${path} >>mergelist_vbm2


```

#step 2e:Averaged image is flipped along x-axis and the two mirror images then re-averaged --> GM template 
```{bash, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
#!/bin/bash
#SBATCH --partition=bigmem
#SBATCH --job-name=fsl
#SBATCH --time=01:00:00
#SBATCH --mem=4G
#SBATCH --chdir=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl
#SBATCH --ntasks=1
#SBATCH --output=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/template1.log
#SBATCH --error=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/template1.err
#SBATCH --cpus-per-task=1

tps=$(date)
echo $tps

module load FSL

wd='/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl'
ukb="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
fsl=${ukb}/FSLVBM
fsl_f=${ukb}/FSLVBM_files

cd ${fsl_f}/Template_anat
mergelist=`cat mergelist_anat_2`
$FSLDIR/bin/fslmerge -t template_4D_GM $mergelist
$FSLDIR/bin/fslmaths template_4D_GM -Tmean template_GM
$FSLDIR/bin/fslswapdim template_GM -x y z template_GM_flipped
$FSLDIR/bin/fslmaths template_GM -add template_GM_flipped -div 2 template_GM

cd ${fsl_f}/Template_vbm
mergelist=`cat mergelist_vbm_2`
$FSLDIR/bin/fslmerge -t template_4D_GM $mergelist
$FSLDIR/bin/fslmaths template_4D_GM -Tmean template_GM
$FSLDIR/bin/fslswapdim template_GM -x y z template_GM_flipped
$FSLDIR/bin/fslmaths template_GM -add template_GM_flipped -div 2 template_GM


tps_fin=$(date)
echo $tps_fin
```

#step 3a: Register all images to template
```{bash, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
#!/bin/bash
#SBATCH --partition=bigmem
#SBATCH --job-name=fsl
#SBATCH --time=00:25:00
#SBATCH --mem=1G
#SBATCH --chdir=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl
#SBATCH --ntasks=1
#SBATCH --array=1-200
#SBATCH --output=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/fsl.log
#SBATCH --error=/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/logs/fsl.err
#SBATCH --cpus-per-task=1

module load FSL

wd='/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl'
ukb="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
fsl=${ukb}/FSLVBM
fsl_f=${ukb}/FSLVBM_files
template_GM_anat=${fsl_f}/Template_anat/template_GM
template_GM_vbm=${fsl_f}/Template_vbm/template_GM

cd ${wd}
export participants=id.csv
dat_0=$(sed -n ${SLURM_ARRAY_TASK_ID}p $participants)

dat=`echo ${dat_0}| xargs`

subject=$(echo $dat | cut -d'_' -f 1)

fslsub=${fsl}/${subject}

#fslvbm3a for pipeline anat 
cd ${fslsub}'/anat'
for g in `$FSLDIR/bin/imglob *_struc.*`
do 
${FSLDIR}/bin/fsl_reg ${g} $template_GM_anat ${g}_GM_to_template_GM -fnirt "--config=GM_2_MNI152GM_2mm.cnf --jout=${g}_JAC_nl"
$FSLDIR/bin/fslmaths ${g}_GM_to_template_GM -mul ${g}_JAC_nl ${g}_GM_to_template_GM_mod -odt float
done 

#fslvbm3a for pipeline vbm 

cd ${fslsub}'/vbm'
cd struc 
for g in `$FSLDIR/bin/imglob *_struc.*`
do 
${FSLDIR}/bin/fsl_reg ${g}_GM $template_GM_vbm ${g}_GM_to_template_GM -fnirt "--config=GM_2_MNI152GM_2mm.cnf --jout=${g}_JAC_nl"
$FSLDIR/bin/fslmaths ${g}_GM_to_template_GM -mul ${g}_JAC_nl ${g}_GM_to_template_GM_mod -odt float
done 
```

#create a mask on all output files, that represents voxels that are equal to zero for all output images 
```{bash, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
vbm="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/FSLVBM"
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"

cd ${wd}
bind="/network/lustre/iss02/ukbiobank/software" 
${bind}/qsubshcom "source /network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/code/source_conda|;
conda activate r_env|;
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"|;
vbm="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/FSLVBM"|;
Rscript /network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/code/mask_zero.R Downloadfsl/participants.tsv \${wd} \${vbm} "anat/T1_brain_pve_1_struc_GM_to_template_GM_mod.nii.gz"|;  
" 1 12G fsl_mask 10:00:00 " "
```



#add script that keep only gm mod registered nifti, and convert it to tsv 
```{bash, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/lustre/iss02/ukbiobank/software" 
cd ${wd}
${bind}/qsubshcom "dat=\$(sed -n \"\${TASK_ID}{p;q}\" ${wd}/Downloadfsl/participants.tsv)|;
mkdir -p ${wd}/FSLVBM/\${dat}/output|;
mkdir -p ${wd}/FSLVBM/\${dat}/output/vbm|;
mkdir -p ${wd}/FSLVBM/\${dat}/output/anat|;
cp ${wd}/FSLVBM/\${dat}/vbm/T1_brain.nii.gz ${wd}/FSLVBM/\${dat}/output/T1_brain.nii.gz|;
cp ${wd}/FSLVBM/\${dat}/vbm/struc/T1_brain_struc_GM_to_template_GM_mod.nii.gz ${wd}/FSLVBM/\${dat}/output/vbm/T1_brain_struc_GM_to_template_GM_mod.nii.gz|;
cp ${wd}/FSLVBM/\${dat}/vbm/struc/T1_brain_struc_GM_to_template_GM.nii.gz ${wd}/FSLVBM/\${dat}/output/vbm/T1_brain_struc_GM_to_template_GM.nii.gz|;
cp ${wd}/FSLVBM/\${dat}/anat/T1_brain_pve_1_struc_GM_to_template_GM_mod.nii.gz ${wd}/FSLVBM/\${dat}/output/anat/T1_brain_pve_1_struc_GM_to_template_GM_mod.nii.gz|;
cp ${wd}/FSLVBM/\${dat}/anat/T1_brain_pve_1_struc_GM_to_template_GM.nii.gz ${wd}/FSLVBM/\${dat}/output/anat/T1_brain_pve_1_struc_GM_to_template_GM.nii.gz|;
source /network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/code/source_conda|;
conda activate r_env|;
Rscript /network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/code/nii_to_tsv.R  ${wd}/FSLVBM/\${dat} ${wd}/FSLVBM_files/Masks/mask_zero_vbm_mod.nii.gz ${wd}/FSLVBM_files/Masks/mask_zero_anat_mod.nii.gz ${wd}/FSLVBM/\${dat}/output|;
rm -r ${wd}/FSLVBM/\${dat}/anat|;
rm -r ${wd}/FSLVBM/\${dat}/vbm|;
" 1 4G nii_to_tsv 00:30:00 "-array=1-42789" 
```


#corresponding Rscript 
```{R, echo=FALSE, message=FALSE, warning=FALSE,eval=F}
arg = commandArgs(trailingOnly=TRUE)

#entrée liste des paths 
library(oro.nifti)

wd=arg[1]
id=arg[2]
mask_dir_vbm=arg[3]
mask_dir_anat=arg[4]
output_dir=arg[5]

filedir=paste(wd,id,sep='/')
img_vbm<-readNIfTI(paste(filedir,'vbm/struc/T1_brain_struc_GM_to_template_GM_mod.nii.gz',sep='/'),reorient=FALSE)
img_anat<-readNIfTI(paste(filedir,'anat/T1_brain_pve_1_struc_GM_to_template_GM_mod.nii.gz',sep='/'),reorient=FALSE)
mask_vbm<-readNIfTI(mask_dir_vbm,reorient=FALSE)
mask_anat<-readNIfTI(mask_dir_anat,reorient=FALSE)
c_mask_vbm<-c(mask_vbm)
c_mask_anat<-c(mask_anat)

c_vbm<-c(img_vbm)
c_anat<-c(img_anat)

c_vbm<-c_vbm[c_mask_vbm!=0]
c_anat<-c_anat[c_mask_anat!=0]

write.table(c_vbm,paste(ouput_dir,'struc/vbm_mod_masked.tsv',sep='/'),row.names=F,col.names=F)
write.table(c_anat,paste(ouput_dir,'anat/anat_mod_masked.tsv',sep='/'),row.names=F,col.names=F)


```

#add ID to anat and vbm tsv
```{bash, message=FALSE, eval=F}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/lustre/iss02/ukbiobank/software" 
cd ${wd}
${bind}/qsubshcom "dat=\$(sed -n \"\${TASK_ID}{p;q}\" ${wd}/Downloadfsl/participants.tsv)|;
cd ${wd}/FSLVBM/\${dat}/output/anat|;
echo \${dat} > anat_mod_masked_id.tsv|;
cat anat_mod_masked.tsv>>anat_mod_masked_id.tsv|;
rm anat_mod_masked.tsv|;
cd ${wd}/FSLVBM/\${dat}/output/vbm|;
echo \${dat} > vbm_mod_masked_id.tsv|;
cat vbm_mod_masked.tsv>>vbm_mod_masked_id.tsv|;
rm vbm_mod_masked.tsv|;
" 1 4G id_to_tsv 00:30:00 "-array=2-42789" 

```

#create file with all output .tsv from each pipeline
#a batch = 500 participants
```{bash, message=FALSE, eval=F}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
dl="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl"
bind="/network/lustre/iss02/ukbiobank/software" 
cd ${wd}/FSLVBM
for batch in {01..86}
do 
echo>tsv_anat_batch${batch}
for elem in `cat ${dl}/participants.batch${batch}`
do 
paste tsv_anat_batch${batch} $elem/output/anat/anat_mod_masked_id.tsv>anat_tmp_batch${batch}
cp anat_tmp_batch${batch} tsv_anat_batch${batch}
done 
rm anat_tmp_batch${batch}
done 

wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
dl="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/Downloadfsl"
bind="/network/lustre/iss02/ukbiobank/software" 
cd ${wd}/FSLVBM
for batch in {01..86}
do 
echo>tsv_vbm_batch${batch}
for elem in `cat ${dl}/participants.batch${batch}`
do 
paste tsv_vbm_batch${batch} $elem/output/vbm/vbm_mod_masked_id.tsv>vbm_tmp_batch${batch}
cp vbm_tmp_batch${batch} tsv_vbm_batch${batch}
done 
rm vbm_tmp_batch${batch}
done 




```



#Merging all batches 
#create BodFiles 

# Import text files in OSCA

```{bash, message=FALSE, eval=F}

wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/lustre/iss02/ukbiobank/software" 


mkdir -p ${wd}/FSLVBM/BodFilesFSL
od="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/FSLVBM"
cd ${od}
proc=vbm
for batch in {01..86}
do
# Cortical
${bind}/qsubshcom " paste voxel_${proc} tsv_${proc}_batch${batch}>tsv_${proc}_batch${batch}vox|;
rm tsv_${proc}_batch${batch}|;
${bind}/osca --tefile ${od}/tsv_${proc}_batch${batch}vox --make-bod --no-fid --out ${od}/BodFilesFSL/${proc}.batch${batch}" 1 10G ${proc}_MakeBod${batch} 10:00:00 "" 
done


```

#Merge all bod into one 
#mybod.flist is a list with paths to bod files to be merged
```{bash, message=FALSE, eval=F}

wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/lustre/iss02/ukbiobank/software" 

cd ${wd}/BodFilesFSL
${bind}/qsubshcom " "${bind}"/osca --befile-flist "${wd}"/FSLVBM/BodFilesFSL/mybod.flistvbm --make-bod --no-fid --out "${wd}"/BodFilesFSL/UKB_VbmFinalBod " 1 8G Merging_Bod 02:00:00 "" 

${bind}/qsubshcom " "${bind}"/osca --befile-flist "${wd}"/FSLVBM/BodFilesFSL/mybod.flistanat --make-bod --no-fid --out "${wd}"/BodFilesFSL/UKB_AnatFinalBod " 1 8G Merging_Bod 02:00:00 "" 
```


#extracting mean and var from BodFiles
```{bash, message=FALSE, eval=F}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/lustre/iss02/ukbiobank/software" 

cd ${wd}/BodFilesFSL


#mean and var for total bod 
${bind}/qsubshcom " "${bind}"/osca --befile "${wd}"/BodFilesFSL/UKB_AnatFinalBod --get-mean --get-variance --out "${wd}"/BodFilesFSL/fsl_anat " 4 100G fsl_var_anat2 48:00:00 "" 

${bind}/qsubshcom " "${bind}"/osca --befile "${wd}"/BodFilesFSL/UKB_VbmFinalBod  --get-mean --get-variance --out "${wd}"/BodFilesFSL/fsl_vbm " 4 100G fsl_var_vbm 48:00:00 "" 

```

#we want to keep only best voxels (ie the one that represent GM), we find a threshold according to mean and var of each vox 
#plotting mean and var for each processing

```{R, message=FALSE, eval=F}

mean_anat<-read.table('../fsl_anat.mean.txt')
var_anat<-read.table('../fsl_anat.var.txt')

df<-data.frame(Mean=mean_vbm$V2,Var=var_vbm$V2,vox=mean_vbm$V1)
ggplot(df,aes(Mean,Var))+geom_point(size=0.01)+ggtitle('FSLVBM')+theme(axis.text.x=element_text(size=16,face="bold"),axis.text.y=element_text(size=16,face="bold"),axis.title = element_text(size=24,face='bold'),plot.title=element_text(size=24,face='bold'))

df_cut<-df[df$Mean>0.1 | df$Var>0.01,] #vbm = 181544, anat = 184 637 #threshold on mean and var 

write.table(df_cut$vox,'../vox_to_keep_vbm',row.names=F,col.names=F) #les 181 544 qu'on garde parmi les 248053 pour vbm et 184 637 parmi 248053 pour anat 

#plot these vox on image to get the mask 
#we do the following lines for anat and vbm

library(oro.nifti)
pre_mask<-readNIfTI('../mask_zero_anat_mod.nii.gz',reorient=F)
val<-c(pre_mask)
vox<-seq(1,902629)
pre_mask<-data.frame(val,vox)
ones<-pre_mask[pre_mask$val==1,] #ceux qui valent 1 dans le pré mask, pre mask = on exclut les vox qui sont 0 en mean et var pour tout le monde
corresp_vox<-cbind(ones,data.frame(mask2=seq(1,254019)))
corresp_vox[!(corresp_vox$mask2 %in% df_cut$vox),'val']<-0
mask<-corresp_vox[corresp_vox$val==1,]

dim(mask)

voxels<-seq(1,902629)
img_df<-data.frame(voxels)
img_df$value<-0

img_df[which(img_df$voxels %in% mask$vox),'value']<-1
img_df<-array(img_df$value,dim=c(91,109,91))

writeNIfTI(img_df,file='../test_anat')


#plot for one subject : rm val = 0 , on regarde que val <0.1 et on plot hist pour les 3 process
im_cat12<-read.table('../../Filezilla_dl/mwp1T1_brain_id_3217683.tsv')
im_vbm<-c(readNIfTI('../../Filezilla_dl/T1_brain_struc_GM_to_template_GM_mod_3217683.nii.gz'))
im_anat<-c(readNIfTI('../../Filezilla_dl/T1_brain_pve_1_struc_GM_to_template_GM_mod_3217683.nii.gz'))

im_cat12<-im_cat12$V1[-1]
im_cat12<-im_cat12[im_cat12!=0] #345 207
im_vbm<-im_vbm[im_vbm!=0] #160 230
im_anat<-im_anat[im_anat!=0] #164 565

df_cat12<-data.frame(Val=im_cat12,Process='CAT12')
df_vbm<-data.frame(Val=im_vbm,Process='FSLVBM')
df_anat<-data.frame(Val=im_anat,Process='FSLANAT')

df_cat12<-df_cat12[df_cat12$Val<0.02,] #167 358
df_vbm<-df_vbm[df_vbm$Val<0.02,] # 22 634
df_anat<-df_anat[df_anat$Val<0.02,] # 22 266

df<-rbind(df_cat12,df_vbm,df_anat)
ggplot(df,aes(x=Val))+
geom_histogram(data=df[df$Process=='FSLVBM',],fill='blue',alpha=0.2,bins=300)+
geom_histogram(data=df[df$Process=='FSLANAT',],fill='green',alpha=0.2,bins=300)+
geom_histogram(data=df[df$Process=='CAT12',],fill='red',alpha=0.2,bins=300)


```



#Get only voxels under cut off (cut_off.R file)

```{bash, message=FALSE, eval=F}

wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/lustre/iss02/ukbiobank/software" 

cd ${wd}/BodFilesFSL

for fsl in Anat Vbm 
do
${bind}/qsubshcom " "${bind}"/osca --befile "${wd}"/BodFilesFSL/UKB_${fsl}FinalBod --extract-probe vox_to_keep_${fsl} --make-bod --no-fid --out "${wd}"/BodFilesFSL/UKB_${fsl}_SubVox " 4 128G sub_Bod_${fsl} 48:00:00 "--partition=bigmem" 
done 

```


#dividing into main and replicate 
```{bash, message=FALSE, eval=FALSE}

wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/lustre/iss02/ukbiobank/software" 

cd ${wd}/BodFilesFSL

for fsl in Anat Vbm
do 
for div in replicate main
do
${bind}/qsubshcom " "${bind}"/osca --befile "${wd}"/BodFilesFSL/UKB_${fsl}_SubVox --keep "${wd}"/BodFilesFSL/${div}_id.csv  --make-bod --no-fid --out "${wd}"/BodFilesFSL/UKB_${fsl}.FinalBod.${div} " 4 128G Bod_${fsl}.${div} 48:00:00 "--partition=bigmem" 
done 
done 
```



#BRM
```{bash, message=FALSE,eval=FALSE}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/lustre/iss02/ukbiobank/software" 

cd ${wd}/BodFilesFSL

for fsl in Anat Vbm
do 
for div in replicate main
do
${bind}/qsubshcom " "${bind}"/osca --befile "${wd}"/BodFilesFSL/UKB_${fsl}.FinalBod.${div} --make-orm-bin --out "${wd}"/BodFilesFSL/UKB_${fsl}.FinalBod.${div}.BRM --thread-num 5 " 8 120G FSL_BRM_calculation 10:00:00 "--partition=bigmem" 
done 
done 

```


#compare both BRM 

```{R, message=FALSE, eval=FALSE}
library(png)
library(ggplot2)

source("RR_4.0_BRM_QC_functions.R")

brm1<-asBRM(readORMBin('UKB_Vbm.FinalBod.main.BRM'))
brm2<-asBRM(readORMBin('UKB_Anat.FinalBod.main.BRM'))

cor_fin<c()
for (i in 1:dim(brm1)[1]){
  cor_fin<-c(cor_fin,cor(brm1[i,],brm2[i,]))
}

cor<-data.frame(cor=cor_fin)
cor$id<-seq(1,dim(brm1)[1])

png("corr_anat_vbm.png", width = 60, height = 30, units = "cm", res = 400)
par(mfrow=c(1,2))
ggplot(corr,aes(x=id,y=cor))+geom_point(size=0.01)
dev.off()

png("corr_anat_vbm2.png", width = 60, height = 30, units = "cm", res = 400)
par(mfrow=c(1,2))
ggplot(corr,aes(x=cor))+geom_histogram(binwidth = 0.01,fill="#69b3a2", color="#e9ecef", alpha=0.9)
dev.off()


```

#final
```{R}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```





