---
title: "Phenotype for LMM"
author: "by Elise Delzant"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
   epuRate::PCTG:
    toc: TRUE
    number_sections: FALSE
    code_folding: "show"
---



```{r, echo=FALSE}
options(width = 60)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

```

#columns to extract
```{R,message=FALSE,eval=F}
library(xlsx)
xl<-read.xlsx(file = '/Users/elise.delzant/Documents/01GM_Association/Phenotypes/ukb_pheno.xlsx',sheetIndex = 1)
df_xl<-as.data.frame(xl)
extract_feat<-df_xl$UKB_extract_field
chr_feat<-c(1)
for (elem in extract_feat){
  chr_feat<-paste(chr_feat,elem,sep=',')
}
chr_feat
```


#extract .tab from ukb with our ID of interest 
```{bash, include=FALSE,eval=F}
wd="/network/iss/ukbiobank/unorganized/unorganized_pheno"
od="/network/iss/ukbiobank/clinic/pheno/organized_pheno"
bind="/network/iss/ukbiobank/software" 
cd $od

${bind}/qsubshcom "cut -f1,7,8,57,58,59,61,62,63,73,76,80,471,472,473,474,475,476,477,478,479,739,740,741,862,863,864,874,875,876,878,879,880,1064,1065,1066,1159,1160,1161,1163,1164,1165,1175,1176,1177,1265,1266,1267,1273,1274,1275,1337,1338,1339,1345,1346,1347,1393,1394,1395,1413,1414,1415,1417,1418,1419,1421,1422,1423,1580,1581,1582,1584,1585,1586,1624,1625,1626,1679,1680,1681,1691,1692,1693,1699,1700,1701,1771,1772,1773,3971,3972,3973,3975,3976,3977,3979,3980,3981,3983,3984,3985,3987,3988,3989,3991,3992,3993,4019,4020,4021,5293,5294,5295,5297,5298,5299,5361,5362,5363,5365,5366,5367,7223,7224,7225,7226,7227,7228,7229,7230,7231,7232,7233,7234,7271,7272,7273,7274,7275,7276,7277,7278,7279,7280,7281,7282,7283,7284,7285,7286,7287,7288,7289,7290,7291,7520,7530,7531,9401,9402,9403,9424,9427,9430,9433,10244,10245,10375,10377,10380,10908,10918,10923,10933,10966,10970,10974,11024,11106,11107,11108,11116,11118,11119,13330,13331,13352,14778,14780,14784,14798,14828,14830,14832,15141,15153,15178,15214,15284,15290,15346,15352,15620,15822,20707,20721,20733 ${wd}/ukb673035.tab > ${od}/ukb673035_UKbiobank_ED_OscaPhenotype.tab|;
" 1 10G UKB_pheno_extract 24:00:00 " "

```

#extract ethnicity
```{bash,eval=F}
wd="/network/iss/ukbiobank/unorganized/unorganized_pheno"
od="/network/iss/ukbiobank/clinic/pheno/organized_pheno"
bind="/network/iss/ukbiobank/software" 
cd $od

${bind}/qsubshcom "cut -f1,1102,11103,11104,11105 ${wd}/ukb673035.tab > ${od}/ukb673035_UKbiobank_ED_Ethnicity.tab|;
" 1 10G UKB_pheno_extract 24:00:00 " "
```

#keep only subjects that have a T1 and separate into T1 20252 and T1 20263, and participants who want to be rm 
```{r,message=FALSE,echo=FALSE,eval=F}
pheno_init<-read.table("/Users/elise.delzant/Documents/01GM_Association/ukb673035_UKbiobank_ED_OscaPhenotype.tab",header=FALSE)
rm_part<-read.table('/Users/elise.delzant/Documents/01GM_Association/UKB_rm_list.csv',header=F)
pheno<-pheno_init
colnames(pheno)=c('ID',as.list(df_xl$Column_name_r))
#keep ID that have 20252 and 20263
pheno_T1_nifti<-pheno[!is.na(pheno$T1_nifti),]
pheno_T1_FS<-pheno[!is.na(pheno$T1_FreeSurfer),]


pheno_T1_nifti<-pheno_T1_nifti[-1,]
pheno_T1_FS<-pheno_T1_FS[-1,]

unusable_20252<-read.table("/Users/elise.delzant/Documents/01GM_Association/unusable_20252.csv",header=FALSE)
pheno_T1_nifti=pheno_T1_nifti[!pheno_T1_nifti$ID %in% unusable_20252$V1,]

unusable_20263<-read.table('/Users/elise.delzant/Documents/01GM_Association/final_id_FS.csv',header=F)
pheno_T1_FS<-pheno_T1_FS[pheno_T1_FS$ID %in% unusable_20263$V1,]
dim(pheno_T1_FS)
dim(pheno_T1_nifti)
pheno<-pheno[(!(is.na(pheno$T1_nifti)) | !is.na(pheno$T1_FreeSurfer)), ]
pheno<-pheno[-1,]


pheno<-pheno[pheno$ID %in% pheno_T1_FS$ID & pheno$ID %in% pheno_T1_nifti$ID,]
pheno<-pheno[pheno$T2_Flair==1 & !is.na(pheno$T2_Flair),]
pheno_no_T2<-pheno[pheno$T2_Flair==0 & !is.na(pheno$T2_Flair),]

#remove participants who requested not to be used
pheno<-pheno[!pheno$ID %in% rm_part$V1,]

dim(pheno)
```

#birth date and year
```{r,message=FALSE,eval=F}
#Field 31
library(dplyr)
library(lubridate)
pheno$Sexe<-as.factor(pheno$Sexe)
levels(pheno$Sexe)<-c('Female','Male')

#Field 11093 
pheno$Assessment_center_age<-as.numeric(pheno$Assessment_center_age)

#Add birth_date 
pheno$Birth_year<-as.numeric(pheno$Birth_year)
pheno$Birth_month<-as.numeric(pheno$Birth_month)
pheno$Birth_date<-make_date(year=pheno$Birth_year,month = pheno$Birth_month,day = 1)


```

#Create Total Brain Volume feature
#Fields 26556 26587 26553 26584
```{r,message=FALSE,eval=F}
missings<-pheno[is.na(pheno$GMV) | is.na(pheno$Cerebral_WMV_l) | is.na(pheno$Cerebral_WMV_r) | is.na(pheno$Cerebell_WMV_l) | is.na(pheno$Cerebell_WMV_r),]
pheno<-pheno[!(pheno$ID %in% missings$ID),]
pheno$TBV<-as.numeric(pheno$GMV)+as.numeric(pheno$Cerebral_WMV_l)+as.numeric(pheno$Cerebral_WMV_r)+as.numeric(pheno$Cerebell_WMV_l)+as.numeric(pheno$Cerebell_WMV_r)



```

#Depression and bipolar status with field 20126

```{r,message=FALSE,eval=F}
pheno$Dep_status<-as.factor(pheno$Dep_status)
levels(pheno$Dep_status)<-c('No Bipolar or depression','Bipolar I','Bipolar II','Major depression (severe)','Major depression (moderate)','Single major depression episod')
pheno$Depression<-as.factor(ifelse(pheno$Dep_status %in% c('No Bipolar or depression','Bipolar I','Bipolar II'),'No depression',pheno$Dep_status))
levels(pheno$Depression)<-c('Major depression (severe)','Major depression (moderate)','Single major depression episod','No depression')
```


#Excluding former drinkers from frequency drinking alcohol  
```{r,message=FALSE,eval=F}
pheno[which(pheno$Alcohol_freq=='-818'),'Alcohol_freq']<-NA

pheno[which(is.na(pheno$Alcohol_ever_dependent)),'Alcohol_ever_dependent']<-2 #new value that indicates NA stands for not ever having been addicted 
pheno[which(pheno$Alcohol_ever_dependent=='-818'),'Alcohol_freq']<-NA
pheno[which(pheno$Alcohol_ever_dependent=='-121'),'Alcohol_freq']<-NA

pheno[pheno$Alcohol_ever_dependent==1 & ! is.na(pheno$Alcohol_ever_dependent),"Alcohol_freq"]<-NA #in Alcohol freq we want only people that have not been formerly addicted 
pheno$Alcohol_freq<-as.factor(pheno$Alcohol_freq)
levels(pheno$Alcohol_freq)<-c('Never','Monthly or less','2-4 t./month','2-3t./week','4+t./week')
```

#stroke, alzheimer and parkinson
```{r,message=FALSE,eval=F}
pheno$Stroke<-as.factor(ifelse(is.na(pheno$Stroke_date),"No","Yes"))
pheno$Stroke_age<-as.numeric(as.Date(pheno$Stroke_date)-pheno$Birth_date)/365.25
#5 subjects with Stroke date 01/01/1900
pheno$Alzheimer<-as.factor(ifelse(is.na(pheno$Alzheimer_date),"No","Yes"))
pheno$Alzheimer_age<-as.numeric(as.Date(pheno$Alzheimer_date)-pheno$Birth_date)/365.25
pheno$Parkinson<-as.factor(ifelse(is.na(pheno$Parkinson_date),"No","Yes"))
pheno$Parkinson_age<-as.numeric(as.Date(pheno$Parkinson_date)-pheno$Birth_date)/365.25


#get infos on when alzheimer, how many years after imaging
alz<-pheno[pheno$Alzheimer=='Yes',]
alz<-alz[,c('Alzheimer_date','Date_center_V2')]
alz$Diff<-as.Date(alz$Alzheimer_date)-as.Date(alz$Date_center_V2)
summary(as.numeric(alz$Diff))

#get infos on when Parkinson, how many years after imaging
alz<-pheno[pheno$Parkinson=='Yes',]
alz<-alz[,c('Parkinson_date','Date_center_V2')]
alz$Diff<-as.Date(alz$Parkinson_date)-as.Date(alz$Date_center_V2)
summary(as.numeric(alz$Diff))
```

#merging visit 0, 1 and 2 
```{r,message=FALSE,eval=F}
#Field 21001
pheno$BMI<-as.numeric(ifelse(is.na(pheno$BMI_V2),ifelse(is.na(pheno$BMI_V1),pheno$BMI_V0,pheno$BMI_V1),pheno$BMI_V2))

#Field 845 value -2 equals never went to school 
pheno[which(pheno$Education_age_V2%in%c('-1','-3')),'Education_age_V2']<-NA
pheno[which(pheno$Education_age_V1%in%c('-1','-3')),'Education_age_V1']<-NA
pheno[which(pheno$Education_age_V0%in%c('-1','-3')),'Education_age_V0']<-NA
pheno$Education_age<-as.numeric(ifelse(is.na(pheno$Education_age_V2),ifelse(is.na(pheno$Education_age_V1),pheno$Education_age_V0,pheno$Education_age_V1),pheno$Education_age_V2))
pheno[which(pheno$Education_age==-2),'Education_age']<-0
#Field 1200
pheno[which(pheno$Sleeplessness_V2=='-3'),'Sleeplessness_V2']<-NA
pheno[which(pheno$Sleeplessness_V1=='-3'),'Sleeplessness_V1']<-NA
pheno[which(pheno$Sleeplessness_V0=='-3'),'Sleeplessness_V0']<-NA
pheno$Sleeplessness<-as.factor(ifelse(is.na(pheno$Sleeplessness_V2),ifelse(is.na(pheno$Sleeplessness_V1),pheno$Sleeplessness_V0,pheno$Sleeplessness_V1),pheno$Sleeplessness_V2))

levels(pheno$Sleeplessness)<-c('Never/rarely','Sometimes','Usually')

#1 Never/rarely
#2 Sometimes
#3 Usually

#set "prefer not to answer to NA
for(i in 0:2){
    var_name <- paste0("Tobacco_current_V",i)
    pheno[which(pheno[,var_name] =="-3"),var_name] <- NA
  
    var_name <- paste0("Tobacco_past_V",i)
    pheno[which(pheno[,var_name] =="-3"),var_name] <- NA
}

#Field 1239 and 1249 
#new var smoking 
pheno$smoking<-NA
pheno[which(pheno$Tobacco_past_V0=="2" | pheno$Tobacco_past_V0=="4" |  pheno$Tobacco_past_V0=="3" & pheno$Tobacco_past_V1=="2" | pheno$Tobacco_past_V1=="4" |  pheno$Tobacco_past_V1=="3"  & pheno$Tobacco_past_V2=="2" | pheno$Tobacco_past_V2=="4" |  pheno$Tobacco_past_V2=="3" & pheno$Tobacco_current_V0=="1" | pheno$Tobacco_current_V0=="0" & pheno$Tobacco_current_V1=="1" | pheno$Tobacco_current_V1=="0" & pheno$Tobacco_current_V2=="1" | pheno$Tobacco_current_V2=="0"),"smoking"]<-"Occasionnally"

pheno[which(pheno$Tobacco_current_V0=="0" & pheno$Tobacco_current_V1=="0" & pheno$Tobacco_current_V2=="0" & pheno$Tobacco_past_V0=="4" & pheno$Tobacco_past_V1=="4" & pheno$Tobacco_past_V2=="4"),"smoking"]<-"Never_smoked"
pheno[which(pheno$Tobacco_past_V0=="3" | pheno$Tobacco_past_V1=="3" | pheno$Tobacco_past_V2=="3" & pheno$Tobacco_current_V0=="0"  & pheno$Tobacco_current_V1=="0"  & pheno$Tobacco_current_V1=="0"),"smoking"]<-"Tried"
pheno[which(pheno$Tobacco_past_V0=="1" | pheno$Tobacco_past_V1=="1" | pheno$Tobacco_past_V2=="1" | pheno$Tobacco_current_V0=="2" | pheno$Tobacco_current_V1=="2" | pheno$Tobacco_current_V2=="2"),'smoking']<-"Ever_regularly"
pheno[which(is.na(pheno$smoking) & pheno$Tobacco_current_V2=="1"),'smoking']<-'Occasionnally'
pheno[which(is.na(pheno$smoking) & pheno$Tobacco_past_V2=="3"),'smoking']<-"Tried"

# get an average of smoked cigarettes accross three time points, -10 means less than one cigarette
pheno$Cig_daily_V1<-as.numeric(pheno$Cig_daily_V1)
pheno[which(pheno$Cig_daily_V1==-1),'Cig_daily_V1']<-NA
pheno[which(pheno$Cig_daily_V1==-10),'Cig_daily_V1']<-1

pheno$Cig_daily_V2<-as.numeric(pheno$Cig_daily_V2)
pheno[which(pheno$Cig_daily_V2==-1),'Cig_daily_V2']<-NA
pheno[which(pheno$Cig_daily_V2==-10),'Cig_daily_V2']<-1

pheno$Cig_daily_V3<-as.numeric(pheno$Cig_daily_V3)
pheno[which(pheno$Cig_daily_V3==-1),'Cig_daily_V3']<-NA
pheno[which(pheno$Cig_daily_V3==-10),'Cig_daily_V3']<-1


pheno$cigarettes_daily<-rowMeans(pheno[,c('Cig_daily_V1','Cig_daily_V2','Cig_daily_V3')])


#check that non smokers and "tried" have not answer this question 

if(nrow(pheno[which(pheno$smoking == "Tried" & !is.na(pheno$cigarettes_daily)),]) !=0){print("We have someone who's only tried a cigarette once or twice who indicates amounts of cigarettes smoked"); break}

if(nrow(pheno[which(pheno$smoking == "Never_smoked" & !is.na(pheno$cigarettes_daily)),]) !=0){print("We have someone who's never smoked who indicates amounts of cigarettes smoked"); break}


# check that the ones who didn't answer the daily cigarettes question are only occasional smokers 
nrow(pheno[which(pheno$smoking == "Occasionally" & !is.na(pheno$cigarettes_daily)),]) # none either 

# split regular smokers into quantiles based on their daily cigarettes indication
pheno$smoking[which(pheno$cigarettes_daily <= quantile(pheno$cigarettes_daily,0.25, na.rm=T))] <- "Regular_bottomquantile"
pheno$smoking[which(pheno$cigarettes_daily > quantile(pheno$cigarettes_daily,0.25, na.rm=T) & pheno$cigarettes_daily <= quantile(pheno$cigarettes_daily,0.5, na.rm=T))] <- "Regular_2ndquantile"
pheno$smoking[which(pheno$cigarettes_daily > quantile(pheno$cigarettes_daily,0.5, na.rm=T) & pheno$cigarettes_daily <= quantile(pheno$cigarettes_daily,0.75, na.rm=T))] <- "Regular_3rdquantile"
pheno$smoking[which(pheno$cigarettes_daily > quantile(pheno$cigarettes_daily,0.75, na.rm=T) & pheno$cigarettes_daily <= quantile(pheno$cigarettes_daily,1, na.rm=T))] <- "Regular_4thquantile"

# If I define it this way I will have to get rid of the regular smokers who didn't indicate how many cigarettes they use :(
pheno$smoking <- ifelse(pheno$smoking == "Ever_regularly" & is.na(pheno$cigarettes_daily), NA, pheno$smoking)

pheno$smoking<-factor(pheno$smoking,levels=c('Never_smoked','Tried','Occasionnally','Regular_bottomquantile','Regular_2ndquantile','Regular_3rdquantile','Regular_4thquantile'))
#Field 1787
pheno[which(pheno$Maternal_smoking_V2%in%c("-1","-3")),'Maternal_smoking_V2']<-NA
pheno[which(pheno$Maternal_smoking_V1%in%c("-1","-3")),'Maternal_smoking_V1']<-NA
pheno[which(pheno$Maternal_smoking_V0%in%c("-1","-3")),'Maternal_smoking_V0']<-NA

pheno$Maternal_smoking<-as.factor(ifelse(is.na(pheno$Maternal_smoking_V2),ifelse(is.na(pheno$Maternal_smoking_V1),pheno$Maternal_smoking_V0,pheno$Maternal_smoking_V1),pheno$Maternal_smoking_V2))
levels(pheno$Maternal_smoking)<-c('No','Yes')
#Field 2976
pheno[which(pheno$Diabetes_V2%in%c("-1","-3")),'Diabetes_V2']<-NA
pheno[which(pheno$Diabetes_V1%in%c("-1","-3")),'Diabetes_V1']<-NA
pheno[which(pheno$Diabetes_V0%in%c("-1","-3")),'Diabetes_V0']<-NA

pheno$Diabetes<-as.factor(ifelse(is.na(pheno$Diabetes_V2),ifelse(is.na(pheno$Diabetes_V1),pheno$Diabetes_V0,pheno$Diabetes_V1),pheno$Diabetes_V2))
levels(pheno$Diabetes)<-c('No','Yes')
#Field 2966
pheno[which(pheno$High_blood_pressure_age_V2%in%c("-1","-3")),'High_blood_pressure_age_V2']<-NA
pheno[which(pheno$High_blood_pressure_age_V1%in%c("-1","-3")),'High_blood_pressure_age_V1']<-NA
pheno[which(pheno$High_blood_pressure_age_V0%in%c("-1","-3")),'High_blood_pressure_age_V0']<-NA

pheno$High_blood_pressure_age<-as.numeric(ifelse(is.na(pheno$High_blood_pressure_age_V2),ifelse(is.na(pheno$High_blood_pressure_age_V1),pheno$High_blood_pressure_age_V0,pheno$High_blood_pressure_age_V1),pheno$High_blood_pressure_age_V2))

#new variable high blood pressure as category Yes No
pheno$High_blood_pressure<-as.factor(ifelse(is.na(pheno$High_blood_pressure_age),'No','Yes'))

#Field 2976
pheno$Diabetes_age<-as.numeric(ifelse(is.na(pheno$Diabetes_age_V2),ifelse(is.na(pheno$Diabetes_age_V1),pheno$Diabetes_age_V0,pheno$Diabetes_age_V1),pheno$Diabetes_age_V2))

pheno[which(pheno$Diabetes_age %in% c(-1,-3)),'Diabetes_age']<-NA

#Field 4803
pheno[which(pheno$Tinnitus_V2=="-1" |pheno$Tinnitus_V2=="-3"),"Tinnitus_V2"]<-NA
pheno[which(pheno$Tinnitus_V1=="-1" |pheno$Tinnitus_V1=="-3"),"Tinnitus_V1"]<-NA
pheno[which(pheno$Tinnitus_V0=="-1" |pheno$Tinnitus_V0=="-3"),"Tinnitus_V0"]<-NA

pheno$Tinnitus<-as.factor(ifelse(is.na(pheno$Tinnitus_V2),ifelse(is.na(pheno$Tinnitus_V1),pheno$Tinnitus_V0,pheno$Tinnitus_V1),pheno$Tinnitus_V2))
levels(pheno$Tinnitus)<-c(0,1,2,3,4,NA) 
levels(pheno$Tinnitus)<-c('Never','Most or all the time','A lot of the time','Some of the time','Not now but in the past')

#field 2754 age at first birth 
pheno[which(pheno$First_birth_V1 %in% c('-3','-4')),'First_birth_V1']<-NA
pheno[which(pheno$First_birth_V2 %in% c('-3','-4')),'First_birth_V2']<-NA
pheno[which(pheno$First_birth_V3 %in% c('-3','-4')),'First_birth_V3']<-NA

pheno$First_birth_age<-as.numeric(ifelse(is.na(pheno$First_birth_V3),ifelse(is.na(pheno$First_birth_V2),pheno$First_birth_V3,pheno$First_birth_V2),pheno$First_birth_V3))

#Hip circumference
pheno$Hip_circumf<-as.numeric(ifelse(is.na(pheno$Hip_circumf_V2),ifelse(is.na(pheno$Hip_circumf_V1),pheno$Hip_circumf_V0,pheno$Hip_circumf_V1),pheno$Hip_circumf_V2))


#Waist circumference
pheno$Waist_circumf<-as.numeric(ifelse(is.na(pheno$Waist_circumf_V2),ifelse(is.na(pheno$Waist_circumf_V1),pheno$Waist_circumf_V0,pheno$Waist_circumf_V1),pheno$Waist_circumf_V2))

#Nb children (male only)
pheno[which(pheno$Nb_children_V2 %in% c('-1','-3')),'Nb_children_V2']<-NA
pheno[which(pheno$Nb_children_V1 %in% c('-1','-3')),'Nb_children_V1']<-NA
pheno[which(pheno$Nb_children_V0 %in% c('-1','-3')),'Nb_children_V0']<-NA
pheno$Nb_children<-ifelse(is.na(pheno$Nb_children_V2),ifelse(is.na(pheno$Nb_children_V1),pheno$Nb_children_V0,pheno$Nb_children_V1),pheno$Nb_children_V2)

#Nb live birth (female only)
pheno[which(pheno$Nb_live_birth_V2 %in% c('-3')),'Nb_live_birth_V2']<-NA
pheno[which(pheno$Nb_live_birth_V1 %in% c('-3')),'Nb_live_birth_V1']<-NA
pheno[which(pheno$Nb_live_birth_V0 %in% c('-3')),'Nb_live_birth_V0']<-NA
pheno$Nb_live_birth<-ifelse(is.na(pheno$Nb_live_birth_V2),ifelse(is.na(pheno$Nb_live_birth_V1),pheno$Nb_live_birth_V0,pheno$Nb_live_birth_V1),pheno$Nb_live_birth_V2)

#merging both 
pheno$Nb_children_both<-as.factor(ifelse(is.na(pheno$Nb_children),pheno$Nb_live_birth,pheno$Nb_children))

#Age when first sexual intercourse
pheno[which(pheno$Sex_first_age_V2 %in% c('-1','-3')),'Sex_first_age_V2']<-NA
pheno[which(pheno$Sex_first_age_V1 %in% c('-1','-3')),'Sex_first_age_V1']<-NA
pheno[which(pheno$Sex_first_age_V0 %in% c('-1','-3')),'Sex_first_age_V0']<-NA

pheno$Sex_first_age<-as.numeric(ifelse(is.na(pheno$Sex_first_age_V2),ifelse(is.na(pheno$Sex_first_age_V1),pheno$Sex_first_age_V0,pheno$Sex_first_age_V1),pheno$Sex_first_age_V2))

#Scaling T1
pheno$Scaling_T1<-as.factor(pheno$Scaling_T1)

#SNR
pheno$SNR<-as.numeric(pheno$SNR)

#Brain volume
pheno$Brain_volume<-as.numeric(pheno$Brain_volume)

#Discrepancy between T1 and standard template linearly aligned
pheno$Discrepancy_template<-as.numeric(pheno$Discrepancy_template)

```
#cognitiv score
```{r,message=FALSE,eval=F}


#cognitiv score Anna Furtjes 
#Verbal numeric reasoning (Fluid intelligence scores)
VNR_vars<-c('Fluid_intell_score_1_V0','Fluid_intell_score_1_V1','Fluid_intell_score_1_V2','Fluid_intell_score_2')

#Matrix pattern completion 
Matrix_vars<-c('Mat_pat_comp_V2','Mat_pat_comp_V3')

#Symbol digit 
SymbolDigit_vars<-c('Number_digit_2_V2','Number_digit_2_V3','Number_digit_1')

#PairMatching 
PairMatching_vars<-c('Incorrect_match_V01','Incorrect_match_V02','Incorrect_match_V03','Incorrect_match_V11','Incorrect_match_V12','Incorrect_match_V13','Incorrect_match_V21','Incorrect_match_V22','Incorrect_match_V23')

#Reaction time 
RT_vars<-c('Reaction_time_V0','Reaction_time_V1','Reaction_time_V2','Reaction_time_V3')

Tower_vars<-c('Tower_rearr_V2','Tower_rearr_V3')

#Trail Making 
TrailMaking_vars<-c('Duration_alphanumeric_1','Duration_alphanumeric_2')

# item 6350: entry of zero means not completed 
pheno$Duration_alphanumeric_1<-as.numeric(pheno$Duration_alphanumeric_1)
pheno$Duration_alphanumeric_1 <- ifelse(pheno$Duration_alphanumeric_1 <= 0.0, NA, pheno$Duration_alphanumeric_1)

pheno$Duration_alphanumeric_2<-as.numeric(pheno$Duration_alphanumeric_2)
pheno$Reaction_time_V0<-as.numeric(pheno$Reaction_time_V0)
pheno$Reaction_time_V1<-as.numeric(pheno$Reaction_time_V1)
pheno$Reaction_time_V2<-as.numeric(pheno$Reaction_time_V2)
pheno$Reaction_time_V3<-as.numeric(pheno$Reaction_time_V3)
pheno$Incorrect_match_V01<-as.numeric(pheno$Incorrect_match_V01)
pheno$Incorrect_match_V02<-as.numeric(pheno$Incorrect_match_V02)
pheno$Incorrect_match_V03<-as.numeric(pheno$Incorrect_match_V03)
pheno$Incorrect_match_V11<-as.numeric(pheno$Incorrect_match_V11)
pheno$Incorrect_match_V12<-as.numeric(pheno$Incorrect_match_V12)
pheno$Incorrect_match_V13<-as.numeric(pheno$Incorrect_match_V13)
pheno$Incorrect_match_V21<-as.numeric(pheno$Incorrect_match_V21)
pheno$Incorrect_match_V22<-as.numeric(pheno$Incorrect_match_V22)
pheno$Incorrect_match_V23<-as.numeric(pheno$Incorrect_match_V23)
pheno$Fluid_intell_score_1_V0<-as.numeric(pheno$Fluid_intell_score_1_V0)
pheno$Fluid_intell_score_1_V1<-as.numeric(pheno$Fluid_intell_score_1_V1)
pheno$Fluid_intell_score_1_V2<-as.numeric(pheno$Fluid_intell_score_1_V2)
pheno$Fluid_intell_score_2<-as.numeric(pheno$Fluid_intell_score_2)
pheno$Mat_pat_comp_V2<-as.numeric(pheno$Mat_pat_comp_V2)
pheno$Mat_pat_comp_V3<-as.numeric(pheno$Mat_pat_comp_V3)
pheno$Number_digit_1<-as.numeric(pheno$Number_digit_1)
pheno$Number_digit_2_V2<-as.numeric(pheno$Number_digit_2_V2)
pheno$Number_digit_2_V3<-as.numeric(pheno$Number_digit_2_V3)
pheno$Tower_rearr_V2<-as.numeric(pheno$Tower_rearr_V2)
pheno$Tower_rearr_V3<-as.numeric(pheno$Tower_rearr_V3)
all_vars <- c(VNR_vars,Matrix_vars,SymbolDigit_vars,PairMatching_vars,RT_vars,Tower_vars,TrailMaking_vars)
pheno$Tinnitus_yes<-ifelse(is.na(pheno$Tinnitus) | pheno$Tinnitus=='Never',0,1)
#apply reverse function because in some variables a lower value means better cognitiv ability 

reverse=function(a){
    a * (-1)
}


pheno[,c(PairMatching_vars,TrailMaking_vars,RT_vars)]<-apply(pheno[,c(PairMatching_vars,TrailMaking_vars,RT_vars)],2,FUN=reverse)

# Standardise variables
normFunc <- function(x){(x-mean(x, na.rm = T))/sd(x, na.rm = T)}

pheno[,all_vars] <- apply(pheno[,all_vars],2,normFunc)
summary(pheno[,all_vars])

##############################
## 1 Verbal-numerical reasoning
##############################

pheno$mean_VNR <- rowMeans(pheno[,VNR_vars],na.rm=T)

summary(pheno$mean_VNR)

##############################
## 2 Matrix 
##############################

pheno$mean_Matrix <- rowMeans(pheno[,Matrix_vars],na.rm=T)
summary(pheno$mean_Matrix)

##############################
## 3 Symbol Digit
##############################

pheno$mean_SymbolDigit <- rowMeans(pheno[,SymbolDigit_vars],na.rm=T)
summary(pheno$mean_SymbolDigit)

##############################
## 4 Pairs Matching
##############################
pheno$mean_PairMatching <- rowMeans(pheno[,PairMatching_vars],na.rm=T)
summary(pheno$mean_PairMatching)

##############################
## 5 Reaction Time 
##############################
#RT_vars

pheno$mean_RT <- rowMeans(pheno[,RT_vars],na.rm=T)
summary(pheno$mean_RT)


##############################
## 6 Tower 
##############################
# Tower_vars

pheno$mean_Tower <- rowMeans(pheno[,Tower_vars],na.rm=T)

summary(pheno$mean_Tower)

##############################
## 7 TrailMaking
##############################
# TrailMaking_vars
## 0 represents "Trail not completed


pheno$mean_TrailMaking <- rowMeans(pheno[,TrailMaking_vars],na.rm=T)
summary(pheno$mean_TrailMaking)

##################################################
## Create Factor Scores
##################################################
library(lavaan)
gpheno <- pheno[,c("ID","mean_VNR", "mean_Matrix", "mean_SymbolDigit", "mean_PairMatching", "mean_RT", "mean_Tower", "mean_TrailMaking")]

#### remove all participants who have missing values in all cog vars 
if(sum(rowSums(is.na(gpheno)) == 7) == 0){print("Each participant has at least one cognitive measurement - woop")}

# remove outliers for each cognitive variable
colSums(!is.na(gpheno))

for(i in c("mean_VNR", "mean_Matrix", "mean_SymbolDigit", "mean_PairMatching", "mean_RT", "mean_Tower", "mean_TrailMaking")){
    # calculate volume specific mean 
        mean_var<-mean(gpheno[,i],na.rm = T)
    # calculate volume specific sd
        sd_var <- sd(gpheno[,i], na.rm=T)
    # remove values that are beyond 4 SDs
        gpheno[,i][which(gpheno[,i] < mean_var - (sd_var*4) | gpheno[,i] > mean_var + (sd_var*4))]<-NA
}
colSums(!is.na(gpheno))

#build SEM model
gmodel <- "Cog =~ mean_VNR + mean_Matrix + mean_SymbolDigit + mean_PairMatching + mean_RT + mean_Tower + mean_TrailMaking" 

# fit SEM model 
fit<-cfa(gmodel, data = gpheno,missing = "ML")
save(fit, file = "g_score_model_fit.RData")
print(summary(fit))

# predict individual-level factor scores 
test_gpheno <- as.data.frame(lavPredict(fit, newdata = gpheno, type = "lv", append.data = T))

# merge with gpheno to get correct eid's

gpheno <- merge(test_gpheno, gpheno, by=c("mean_VNR", "mean_Matrix", "mean_SymbolDigit", "mean_PairMatching", "mean_RT", "mean_Tower", "mean_TrailMaking"))

gpheno <- gpheno[,c("ID", "Cog")]

# only keep non-missing participants
gpheno <- gpheno[!is.na(gpheno$Cog),]

### realised later OSCA wants IID FID value/ no col.names
gpheno$FID <- gpheno$ID

gpheno <- gpheno[,c("ID","FID","Cog")]

##### remove outliers outside of 4SDs
mean<-mean(gpheno$Cog, na.rm=T)
sd<-sd(gpheno$Cog, na.rm=T)
gpheno$Cog[which(gpheno$Cog < mean - (sd*4) | gpheno$Cog > mean + (sd*4))]<-NA

print(summary(gpheno))

pheno$Cog<-ifelse(pheno$ID %in% gpheno$ID,gpheno$Cog,NA)

```

```{r,message=FALSE,eval=F}
library(lavaan)
library(semPlot)


## fit_reserve gives the same results as fit but with fully positive factor loadings (by reverse I mean that I have multiplied items with negative polarity by -1 so that all loadings are positive)
## we now want to get the variance explained by the model:
## we take the sum of the variance output for each item and divide by the number of items
total_variance_explained <-mean(1-mean(rowSums(inspect(fit,what="std")$theta)))
print(paste0("The total variance explained by this model is ", round(total_variance_explained, digits=2)))

labels <- c("VNR", "Matrix", "Symbol\nDigit", "Pair\nMatching", "RT", "Tower", "Trail\nMaking","Cog")

 semPaths(fit,
                           what = "paths",
                           whatLabels="est", 
                           intercepts=FALSE, 
                           style="OpenMx",
                           layout = "tree",
                           nodeLabels=labels,
                           curvePivot=TRUE, 
                           edge.color="black", 
                           sizeMan = 8,
                           pastel=TRUE, 
                           rainbowStart = 0.5,
                           nDigits = 2, 
                            title = TRUE)

```

#pain
```{r,message=FALSE,eval=F}

#reponses du choix multiple: 
#1 Headache
#2 Facial pain 
#3 Neck or shoulder
#4 Back pain
#5 Stomach pain
#6 Hip pain 
#7 Knee pain 
#8 Pain all over the body 


pheno[which(pheno$Pain_V00 %in% c('-7','-3','8')),'Pain_V00']<-NA
pheno[which(pheno$Pain_V01 %in% c('-7','-3','8')),'Pain_V01']<-NA
pheno[which(pheno$Pain_V02 %in% c('-7','-3','8')),'Pain_V02']<-NA
pheno[which(pheno$Pain_V03 %in% c('-7','-3','8')),'Pain_V03']<-NA
pheno[which(pheno$Pain_V04 %in% c('-7','-3','8')),'Pain_V04']<-NA
pheno[which(pheno$Pain_V05 %in% c('-7','-3','8')),'Pain_V05']<-NA
pheno[which(pheno$Pain_V06 %in% c('-7','-3','8')),'Pain_V06']<-NA
pheno[which(pheno$Pain_V10 %in% c('-7','-3','8')),'Pain_V10']<-NA
pheno[which(pheno$Pain_V11 %in% c('-7','-3','8')),'Pain_V11']<-NA
pheno[which(pheno$Pain_V12 %in% c('-7','-3','8')),'Pain_V12']<-NA
pheno[which(pheno$Pain_V13 %in% c('-7','-3','8')),'Pain_V13']<-NA
pheno[which(pheno$Pain_V14 %in% c('-7','-3','8')),'Pain_V14']<-NA
pheno[which(pheno$Pain_V15 %in% c('-7','-3','8')),'Pain_V15']<-NA
pheno[which(pheno$Pain_V16 %in% c('-7','-3','8')),'Pain_V16']<-NA
pheno[which(pheno$Pain_V20 %in% c('-7','-3','8')),'Pain_V20']<-NA
pheno[which(pheno$Pain_V21 %in% c('-7','-3','8')),'Pain_V21']<-NA
pheno[which(pheno$Pain_V22 %in% c('-7','-3','8')),'Pain_V22']<-NA
pheno[which(pheno$Pain_V23 %in% c('-7','-3','8')),'Pain_V23']<-NA
pheno[which(pheno$Pain_V24 %in% c('-7','-3','8')),'Pain_V24']<-NA
pheno[which(pheno$Pain_V25 %in% c('-7','-3','8')),'Pain_V25']<-NA
pheno[which(pheno$Pain_V26 %in% c('-7','-3','8')),'Pain_V26']<-NA

pheno$all_pains_V0<-0
pheno$all_pains_V1<-0
pheno$all_pains_V2<-0

pain_V0<-function(x){
  all_pains<-c(as.character(x[,c('Pain_V00','Pain_V01','Pain_V02','Pain_V03','Pain_V04','Pain_V05','Pain_V06')]))
  all_pains<-all_pains[!is.na(all_pains)]
  if ('1' %in% all_pains & (x[,c('Headache_V0')]!='1')) all_pains<-all_pains[all_pains!="1"]
  if ('2' %in% all_pains & (x[,c('Facial_pain_V0')]!='1')) all_pains<-all_pains[all_pains!="2"]
  if ('3' %in% all_pains & (x[,c('Neck_pain_V0')]!='1')) all_pains<-all_pains[all_pains!="3"]
  if ('4' %in% all_pains & (x[,c('Back_pain_V0')]!='1')) all_pains<-all_pains[all_pains!="4"]
  if ('5' %in% all_pains & (x[,c('Stomach_pain_V0')]!='1')) all_pains<-all_pains[all_pains!="5"]
  if ('6' %in% all_pains & (x[,c('Hip_pain_V0')]!='1')) all_pains<-all_pains[all_pains!="6"]
  if ('7' %in% all_pains & (x[,c('Knee_pain_V0')]!='1')) all_pains<-all_pains[all_pains!="7"]
  return(length(all_pains))
 
}

pain_V1<-function(x){
  all_pains<-c(as.character(x[,c('Pain_V10','Pain_V11','Pain_V12','Pain_V13','Pain_V14','Pain_V15','Pain_V16')]))
  all_pains<-all_pains[!is.na(all_pains)]
  if ('1' %in% all_pains & (x[,c('Headache_V1')]!='1')) all_pains<-all_pains[all_pains!="1"]
  if ('2' %in% all_pains & (x[,c('Facial_pain_V1')]!='1')) all_pains<-all_pains[all_pains!="2"]
  if ('3' %in% all_pains & (x[,c('Neck_pain_V1')]!='1')) all_pains<-all_pains[all_pains!="3"]
  if ('4' %in% all_pains & (x[,c('Back_pain_V1')]!='1')) all_pains<-all_pains[all_pains!="4"]
  if ('5' %in% all_pains & (x[,c('Stomach_pain_V1')]!='1')) all_pains<-all_pains[all_pains!="5"]
  if ('6' %in% all_pains & (x[,c('Hip_pain_V1')]!='1')) all_pains<-all_pains[all_pains!="6"]
  if ('7' %in% all_pains & (x[,c('Knee_pain_V1')]!='1')) all_pains<-all_pains[all_pains!="7"]
  return(length(all_pains))
 
}

pain_V2<-function(x){
  all_pains<-c(as.character(x[,c('Pain_V20','Pain_V21','Pain_V22','Pain_V23','Pain_V24','Pain_V25','Pain_V26')]))
  all_pains<-all_pains[!is.na(all_pains)]
  if ('1' %in% all_pains & (x[,c('Headache_V2')]!='1')) all_pains<-all_pains[all_pains!="1"]
  if ('2' %in% all_pains & (x[,c('Facial_pain_V2')]!='1')) all_pains<-all_pains[all_pains!="2"]
  if ('3' %in% all_pains & (x[,c('Neck_pain_V2')]!='1')) all_pains<-all_pains[all_pains!="3"]
  if ('4' %in% all_pains & (x[,c('Back_pain_V2')]!='1')) all_pains<-all_pains[all_pains!="4"]
  if ('5' %in% all_pains & (x[,c('Stomach_pain_V2')]!='1')) all_pains<-all_pains[all_pains!="5"]
  if ('6' %in% all_pains & (x[,c('Hip_pain_V2')]!='1')) all_pains<-all_pains[all_pains!="6"]
  if ('7' %in% all_pains & (x[,c('Knee_pain_V2')]!='1')) all_pains<-all_pains[all_pains!="7"]
  return(length(all_pains))
 
}

for (i in seq(1,dim(pheno)[1])){
  x<-pheno[i,]
  pheno[i,'all_pains_V0']<-pain_V0(x)
}

for (i in seq(1,dim(pheno)[1])){
  x<-pheno[i,]
  pheno[i,'all_pains_V1']<-pain_V1(x)
}

for (i in seq(1,dim(pheno)[1])){
  x<-pheno[i,]
  pheno[i,'all_pains_V2']<-pain_V2(x)
}

summary(as.factor(pheno$all_pains_V0))
summary(as.factor(pheno$all_pains_V1))
summary(as.factor(pheno$all_pains_V2))

#merging all into one variable 
pheno$all_pains<-as.factor(ifelse(is.na(pheno$all_pains_V2),ifelse(is.na(pheno$all_pains_V1),pheno$all_pains_V0,pheno$all_pains_V1),pheno$all_pains_V2))


```


#compute mean GM density 
```{r,message=FALSE,eval=F}

arg = commandArgs(trailingOnly=TRUE)

#entrée liste des paths 
library(oro.nifti)
library(scales)
list_path=arg[1]
writing_path=arg[2]
working_directory_fsl=arg[3]
working_directory_cat12=arg[4]

ids<-read.csv(list_path,header = FALSE,sep="")
cat12<-c()
anat<-c()
vbm<-c()
for (im in ids$V1){
  path_fsl<-paste(as.character(working_directory_fsl),as.character(im),sep="/")
  path_cat12<-paste(as.character(working_directory_cat12),as.character(im),sep="/")
  print(paste(path_fsl,"output/vbm/T1_brain_struc_GM_to_template_GM_mod.nii.gz",sep="/"))
  print(paste(path_cat12,"OutputVol/mwp1T1_brain_padded.nii.gz",sep="/"))
  img_i_anat <- c(readNIfTI(paste(path_fsl,"output/vbm/T1_brain_struc_GM_to_template_GM_mod.nii.gz",sep="/"),reorient = FALSE))
  img_i_vbm<-c(readNIfTI(paste(path_fsl,"output/anat/T1_brain_pve_1_struc_GM_to_template_GM_mod.nii.gz",sep="/"),reorient = FALSE))
  img_i_cat12<-c(readNIfTI(paste(path_cat12,"OutputVol/mwp1T1_brain_padded.nii.gz",sep="/"),reorient=FALSE))
  cat12<-c(cat12,mean(c(img_i_cat12)))
  anat<-c(anat,mean(c(img_i_anat)))
  vbm<-c(vbm,mean(c(img_i_vbm)))
}

write.table(data.frame(cat12), file=paste(writing_path,"GM_density_cat12",sep='/'))
write.table(data.frame(anat), file=paste(writing_path,"GM_density_anat",sep='/'))
write.table(data.frame(vbm), file=paste(writing_path,"GM_density_vbm",sep='/'))

```

```{bash,message=FALSE,eval=F}

wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"

cd ${wd}
bind="/network/lustre/iss02/ukbiobank/software" 
${bind}/qsubshcom "source /network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/code/source_conda|;
conda activate r_env|;
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant"|;
fsl="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/FSLVBM"|;
cat12="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/CAT12"|;
Rscript /network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/code/GM_density.R Downloadfsl/participants.tsv \${wd} \${fsl} \${cat12}|;  
" 4 24G mean_GM_density 24:00:00 "--partition=bigmem"

```

#add GM density to pheno df 

```{R,message=FALSE,eval=F}
GM_dens_cat12<-read.table('../../Filezilla_dl/GM_density_cat12',header=F)
GM_dens_vbm<-read.table('../../Filezilla_dl/GM_density_vbm',header=F)
GM_dens_anat<-read.table('../../Filezilla_dl/GM_density_anat',header=F)

participants<-read.table('../../Filezilla_dl/participants.tsv')

df<-data.frame(GM_dens_anat$V1,GM_dens_cat12$V1,GM_dens_vbm$V1,participants$V1)
df<-df[df$participants.V1 %in% pheno$ID,]

pheno$GM_dens_Vol<-df$GM_dens_cat12.V1
pheno$GM_dens_Vbm<-df$GM_dens_vbm.V1
pheno$GM_dens_Anat<-df$GM_dens_anat.V1

```

#total surface area freesurfer val 
```{r,message=FALSE,eval=F}
pheno$Surf_left_hemi<-as.numeric(pheno$Surf_left_hemi)
pheno$Surf_right_hemi<-as.numeric(pheno$Surf_right_hemi)
pheno$Total_surface_area<-pheno$Surf_left_hemi + pheno$Surf_right_hemi

```


#as factors 
```{r,message=FALSE,eval=F}
pheno$Sexe<-as.factor(pheno$Sexe)
levels(pheno$Sexe)<-c('Female','Male')
pheno$Center<-as.factor(pheno$Center)
pheno$T1_nifti<-as.factor(pheno$T1_nifti)
pheno$T1_FreeSurfer<-as.factor(pheno$T1_FreeSurfer)
pheno[which(pheno$Cannabis=="-818"),"Cannabis"]<-NA
pheno$Cannabis<-as.factor(pheno$Cannabis)
levels(pheno$Cannabis)<-c('No','1-2t.','3-10t.','11-100t.','100+')
pheno[which(pheno$Ongoing_addiction=="-818"),"Ongoing_addiction"]<-NA
levels(pheno$Ongoing_addiction)<-c('No','Yes')
pheno$Ongoing_addiction<-as.factor(pheno$Ongoing_addiction)
pheno$First_psychotic_age<-as.factor(pheno$First_psychotic_age)
pheno$Mean_rfmri<-as.numeric(pheno$Mean_rfmri)
pheno$X_pos<-as.numeric(pheno$X_pos)
pheno$Y_pos<-as.numeric(pheno$Y_pos)
pheno$Z_pos<-as.numeric(pheno$Z_pos)
pheno$IMD<-as.numeric(pheno$IMD)
pheno[which(pheno$Restlessness=='-818'),'Restlessness']<-NA
pheno$Restlessness<-as.factor(pheno$Restlessness)
levels(pheno$Restlessness)<-c('Not at all','Several days','More than half the days','Nearly every day')

#Date_centerV2 create gap between first MRI made from center and one's MRI
pheno$Date_center<-as.Date(pheno$Date_center)

pheno_c1<-pheno[pheno$Center==11025,]
pheno_c2<-pheno[pheno$Center==11026,]
pheno_c3<-pheno[pheno$Center==11027,]
pheno_c4<-pheno[pheno$Center==11028,]

min1<-min(pheno_c1$Date_center)
min2<-min(pheno_c2$Date_center)
min3<-min(pheno_c3$Date_center)
min4<-min(pheno_c4$Date_center)

pheno_c1$Diff_to_mri<-as.numeric(pheno_c1$Date_center-min1)
pheno_c2$Diff_to_mri<-as.numeric(pheno_c2$Date_center-min2)
pheno_c3$Diff_to_mri<-as.numeric(pheno_c3$Date_center-min3)
pheno_c4$Diff_to_mri<-as.numeric(pheno_c4$Date_center-min4)

pheno$Diff_to_mri<-NA
pheno[which(pheno$Center==11025),'Diff_to_mri']<-as.numeric(pheno_c1$Date_center-min1)
pheno[which(pheno$Center==11026),'Diff_to_mri']<-as.numeric(pheno_c2$Date_center-min2)
pheno[which(pheno$Center==11027),'Diff_to_mri']<-as.numeric(pheno_c3$Date_center-min3)
pheno[which(pheno$Center==11028),'Diff_to_mri']<-as.numeric(pheno_c4$Date_center-min4)

pheno$Diff_to_mri<-pheno$Diff_to_mri/365.25

save(pheno,file='pheno_plots.RData')
```

#keeping only final variables 
```{r,message=FALSE,eval=F}

#center categories 
cent<-c('ID','Center')

#individual characteristics 
indiv<-c('Sexe','Assessment_center_age','BMI','Education_age','Maternal_smoking','IMD','Sex_first_age','Nb_children_both','Brain_volume','Waist_circumf','Hip_circumf','TBV')

#Addiction
addict<-c('smoking','Alcohol_freq')

#Disease
disease<-c('Sleeplessness','Diabetes','High_blood_pressure','Stroke','Alzheimer','Parkinson','Tinnitus','Restlessness','all_pains')

#Depression and bipolarity
dep<-'Depression'

# 
#T1 related variables
t1<-c('X_pos','Y_pos','Diff_to_mri','Mean_rfmri','Discrepancy_template','Discrepancy_template_non_lin','Scaling_T1','SNR','GM_dens_Vol','GM_dens_Anat','GM_dens_Vbm','Total_surface_area')

#Cognition
cog<-'Cog'

pheno_lmm<-pheno[,c(cent,indiv,addict,disease,dep,t1,cog)]

levels(pheno_lmm$Sexe)<-c(0,1)
levels(pheno_lmm$Maternal_smoking)<-c(0,1)
levels(pheno_lmm$smoking)<-c(0,1,2,3,4,5,6)
levels(pheno_lmm$Alcohol_freq)<-c(0,1,2,3,4)
levels(pheno_lmm$Sleeplessness)<-c(0,1,2)
levels(pheno_lmm$Restlessness)<-c(0,1,2,3)
levels(pheno_lmm$Diabetes)<-c(0,1)
levels(pheno_lmm$Parkinson)<-c(0,1)
levels(pheno_lmm$Stroke)<-c(0,1)
levels(pheno_lmm$High_blood_pressure)<-c(0,1)
levels(pheno_lmm$Tinnitus)<-c(0,1,2,3,4)
levels(pheno_lmm$Depression)<-c(4,3,2,1)
levels(pheno_lmm$Alzheimer)<-c(0,1)
pheno_lmm$Nb_children_both<-as.numeric(pheno_lmm$Nb_children_both)
save(pheno_lmm,file='pheno_lmm.RData')
write.table(colnames(pheno_lmm),'../phenotypes_list',row.names=F,col.names=F)

```

#divide into main and replicate 

```{r,message=FALSE,eval=F}

load('pheno_lmm.RData')

main<-pheno_lmm[which(pheno_lmm$Center==11025),]
replicate<-pheno_lmm[which(pheno_lmm$Center %in% c(11026,11027,11028)),]
save(main,file='../Data/main.Rdata')
save(replicate,file='../Data/replicate.RData')
write.table(as.data.frame(main$ID),file = '/Users/elise.delzant/Documents/01GM_Association/main_id.csv',row.names = FALSE,col.names = FALSE)
write.table(as.data.frame(replicate$ID),file = '/Users/elise.delzant/Documents/01GM_Association/replicate_id.csv',row.names = FALSE,col.names = FALSE)
```


