---
title: "Multiple testing"
author: "by Elise Delzant"
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r, echo=FALSE}
options(width = 60)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

```

Create 100 random traits for each processing and each method (100*6*2)
```{r, message=FALSE, warning=FALSE,eval=F}
load('../01GM_Association/Data/main.Rdata')
load('../01GM_Association/Data/replicate.Rdata')
ids_rm<-read.table('../01GM_Association/BRM/id_QC_rm')
main<-main[!(main$ID %in% ids_rm$V1),]
replicate<-replicate[!(replicate$ID %in% ids_rm$V1),]


set.seed(123)
for (elem in 1:100){
    df<-data.frame(ID1=main$ID,ID2=main$ID)
    samp<-rnorm(23183)
    df$sample<-samp
    df$ID1<-as.factor(df$ID1)
    df$ID2<-as.factor(df$ID2)
    name<-paste('random_pheno','main',elem,sep="_")
    write.table(df,paste('Random_traits',name,sep='/'),col.names=F,row.names=F,quote=FALSE)
    }
  
#we add 400 more other traits 
set.seed(45)
for (elem in 101:500){
    df<-data.frame(ID1=main$ID,ID2=main$ID)
    samp<-rnorm(23183)
    df$sample<-samp
    df$ID1<-as.factor(df$ID1)
    df$ID2<-as.factor(df$ID2)
    name<-paste('random_pheno','main',elem,sep="_")
    write.table(df,paste('Random_traits',name,sep='/'),col.names=F,row.names=F,quote=FALSE)
}

#we add 500 more
set.seed(70)
for (elem in 501:1000){
    df<-data.frame(ID1=main$ID,ID2=main$ID)
    samp<-rnorm(23183)
    df$sample<-samp
    df$ID1<-as.factor(df$ID1)
    df$ID2<-as.factor(df$ID2)
    name<-paste('random_pheno','main',elem,sep="_")
    write.table(df,paste('Random_traits',name,sep='/'),col.names=F,row.names=F,quote=FALSE)
}
```

#First we need to create UKB_QC Bod Files 

```{bash, message=FALSE, warning=FALSE,eval=F}

###FSL
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFSL"
bind="/network/lustre/iss02/ukbiobank/software"
ids="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"

cd ${wd}

for proc in Anat Vbm 
do 
for div in main replicate
do 
${bind}/qsubshcom "${bind}/osca --befile UKB_${proc}.FinalBod.${div} --remove ${ids}/id_QC_rm --make-bod --out UKB_QC_${proc}.FinalBod.${div}|;
" 1 40G ${div}_${proc}_QC_Bod 04:00:00 ""
done
done 

##CAT12
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesCAT12"
bind="/network/lustre/iss02/ukbiobank/software"
ids="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"

cd ${wd}
for proc in Vol Lh.Thick Rh.Thick
do 
for div in main replicate
do 
${bind}/qsubshcom "${bind}/osca --befile UKB_${proc}.FinalBod.${div} --remove ${ids}/id_QC_rm --make-bod --out UKB_QC_${proc}.FinalBod.${div}|;
" 1 40G ${div}_${proc}_QC_Bod 04:00:00 ""
done
done 


###FREESURFER
wd="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant"
bind="/network/iss/ukbiobank/software" 
ids="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"

cd ${wd}/BodFilesFS

for hemi in lh
do
for moda in thickness 
do
for div in replicate main
do
${bind}/qsubshcom " ${bind}/osca --befile ${wd}/BodFilesFS/UKB_FS_${hemi}.${moda}.FinalBod.${div} --remove ${ids}/id_QC_rm --make-bod --out UKB_QC_FS_${hemi}.${moda}.FinalBod.${div}|;
" 1 40G ${div}_${hemi}_${moda}_QC_Bod 04:00:00 ""
done
done 
done



```

#for surface-based processing, merging both hemisphere bod files just created 

#CAT 12 merging BRM for surfaces : careful because Lh Thick and Rh Thick have the same .opi (because they were created separately so both have opi starting with 001, 002 etc)
#when merging, need to modify one .opi by naming probes starting from 163842
#in my case i transformed Lh Hemi vertices value to 16842 16843 ...

#same for freesurfer


###1st : remove vertices to exclude for each hemisphere
###2nd : replacing opis name for one of the hemisphere
###3rd: merge hemisphere



```{bash, message=FALSE, warning=FALSE,eval=F}

wd="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesCAT12"
bind="/network/iss/ukbiobank/software"


cd ${wd}
#need to remove vertices to exclude (list given by baptiste) for CAT12


for div in replicate
do
${bind}/qsubshcom "${bind}/osca --befile UKB_QC_Rh.Thick.FinalBod.${div} --exclude-probe VerticesToExclude_rh_thickness_Cortex.txt --make-bod --out UKB_QC_CAT12_Surface_Rh.FinalBod.${div}|;
" 1 40G ${div}_QC_Bod_Rh 04:00:00 ""
done

for div in replicate 
do
${bind}/qsubshcom "${bind}/osca --befile UKB_QC_Lh.Thick.FinalBod.${div} --exclude-probe VerticesToExclude_lh_thickness_Cortex.txt --make-bod --out UKB_QC_CAT12_Surface_Lh.FinalBod.${div}|;
" 1 40G ${div}_QC_Bod_Lh 04:00:00 ""
done

#rename .opi file dans R
#in rh thick, 149 926 probes. pour lh.thick on additionne +149 926

#merge Lh and Rh hemi
wd="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesCAT12"
bind="/network/iss/ukbiobank/software"
ids="/network//iss/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"

cd ${wd}

for div in replicate
do 
${bind}/qsubshcom "${bind}/osca --befile-flist flist_cat12_surf_${div} --make-bod --out UKB_QC_CAT12_Surface.FinalBod.${div}|;
" 1 40G ${div}_QC_Bod 08:00:00 ""
done





###FREESURFER : first, same as for cat12, need to re create .opi with different probes names
###FREESURFER : merging all modalities for Freesurfer all moda, and merging only right and left hemisphere thickness for FreeSurfer Thickness
wd="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFS"
bind="/network/iss/ukbiobank/software"

cd ${wd}

for div in main replicate
do 
for moda in All_moda
do
${bind}/qsubshcom "${bind}/osca --befile-flist flist_FS_${moda}_${div} --make-bod --out UKB_QC_FS_${moda}.FinalBod.${div}|;
" 1 40G ${div}_QC_Bod 10:00:00 ""
done
done



```

#LR model

```{bash, message=FALSE, warning=FALSE,eval=F}

wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"
bind="/network/lustre/iss02/ukbiobank/software"
fsl="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFSL"
cat12="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesCAT12"
fs="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFS"

cd $wd 

#LINEAR MODEL ok 
#pour Anat_rank_norm --> covar_discr_Anat

for proc in CAT12_Volume
do
for div in main 
do 
for seq in 917 918
do
${bind}/qsubshcom "${bind}/osca --befile ${cat12}/UKB_QC_Vol.FinalBod.main  --pheno Random_traits/random_pheno_${div}_${seq}  --covar covar_discr_${div} --qcovar covar2_quant_Vol_body_size_${div} --linear --out Random_traits/Results/${proc}/resu_${proc}_LR_${seq}|;
" 4 40G ${proc}_${seq} 01:30:00 ""
done 
done 
done 

#We do this for FS_rh.area, FS_lh.area, FS_Thick_LogJacs (previously need to merge lh thick, rh thick, lh logjacs and rh logjacs  bods)
#dont forget to rename probes (.opi) otherwise in rh and lh they have same names for thickness and area
for proc in FS_All_moda_rank_norm
do
for div in main 
do 
for seq in $(seq 1 1 1000)
do
${bind}/qsubshcom "${bind}/osca --befile ${fs}/UKB_QC_${proc}.FinalBod.${div} --pheno Random_traits/random_pheno_${div}_${seq}  --covar covar_discr_${div} --qcovar covar2_quant_fs_body_size_${div} --linear --out Random_traits/Results/${proc}/resu_${proc}_${seq}|;
" 4 150G ${div}_${proc}_${seq} 02:30:00 "--partition=bigmem"
done 
done 
done 



#pour adj probe rank norm 


for proc in FS_All_moda
do
for div in main 
do 
for method in adj_probe.rank_norm
do
for seq in $(seq 501 1 1000)
do
${bind}/qsubshcom "${bind}/osca --befile ${fs}/UKB_QC_${proc}.${method}  --pheno Random_traits/random_pheno_${div}_${seq} --linear --out Random_traits/Results/${proc}.${method}/resu_${proc}_${method}_LR_${seq}|;
" 4 150G ${proc}_${seq} 01:30:00 "--partition=bigmem"
done 
done 
done 
done

```

#for FS, merge FS_thickness FS_lh_area, FS_rh_area, FS_Thick_LogJacs (because osca was not able to merge all bods, so we did the LR analysis separately)

```{R, message=FALSE, warning=FALSE,eval=F}
for (i in seq(441,500)){
  print(i)
  thickness=paste('resu_FS_Thickness_LR_main_',i,'.linear',sep='')
  lh_area=paste('resu_FS_lh.area_LR_main_',i,'.linear',sep='')
  rh_area=paste('resu_FS_rh.area_LR_main_',i,'.linear',sep='')
  logjacs<-paste('resu_FS_Thick_LogJacs_LR_main_',i,'.linear',sep='')
  all_moda<-paste('resu_FS_All_Moda_LR_main_',i,'.linear',sep='')
  th<-read.table(thickness,header=F)
  lh<-read.table(lh_area,header=F)
  rh<-read.table(rh_area,header=F)
  log<-read.table(logjacs,header=F)
  lh<-lh[-c(1),]
  rh<-rh[-c(1),]
  log<-log[-c(1),]
  tab<-rbind(th,lh,rh,log)
  write.table(tab,all_moda,row.names=F,col.names=F)
}

```

#read .linear
#paralleliser 
```{bash, message=FALSE, warning=FALSE,eval=F}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol/Random_traits/Results"
bind="/network/lustre/iss02/ukbiobank/software"
cd ${wd}

for proc in FS_Thickness.adj_probe.rank_norm
do 
for N in 40 50 55 60 65 70 75 80 90 100 105 110 115 120 125 130 135 140 145 150 155 160 165 170 175 180 185 190 195 200 220 240 260 280 184.637 181.544 192.483 299.881 654.002
do
for model in LR
do
${bind}/qsubshcom "module load R/4.1.2|;
Rscript "1_1_Bonferroni.R" ${proc} ${N} ${model} " 1 4G Bonf_${proc}_${N} 02:00:00 ""
done
done
done

#sinon on prend 1_1_Bonferroni.R (la difference= nom du fichier il y a 'main' dedans, alors que pr rank norm non)
#il y a aussi 1_1_Bonferroni2.R --> percent_kurt5
### pour info
#nvox_anat=184637
#nvox_vbm=181544
#nvox_cat12=192483
#nvox_cat12_surf=299881
#nvox_fs=654002
###


```

#to merge all results into one file per processing
```{R, message=FALSE, warning=FALSE,eval=F}
library(data.table)
N=c(50,55,60,65,70,75,80,90,100,105,110,115,120,125,130,135,140,145,150,155,160,165,170,175,180,185,190,195,200,220,240,260,280,184.637,181.544,192.483,299.881,654.002) 


pval=0.05/c(N)
#we do this for all proc. can also make a loop 
proc='FS_Thickness'
path1=paste('pval','40',proc,'LR',sep='_')
df_fin=read.table(path1)
for (n in N){
  path_tab<-paste('pval',as.character(n),proc,'LR',sep='_')
  tab<-read.table(path_tab)
  df_fin=rbind(df_fin,tab)
}
df_fin$N<-c(40,N)*1000
write.table(df_fin,'tab_fin_pval_FS_Thickness_LR_1000',row.names=F,col.names=F)


```

get minimal pvalue per voxel per processing
```{bash, message=FALSE, warning=FALSE,eval=F}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol/Random_traits/Results"
bind="/network/lustre/iss02/ukbiobank/software"
cd ${wd}

 
 
for proc in FS_Thickness_adj_probe.rank_norm
do
for model in LR
do
${bind}/qsubshcom "module load R/4.1.2|;
Rscript "pvalmin.R" ${proc} ${N} ${model} " 4 12G pvalmin_${proc} 04:00:00 "--partition=bigmem"
done
done

```

get minimal pvalue per trait per processing
```{bash, message=FALSE, warning=FALSE,eval=F}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol/Random_traits/Results"
bind="/network/lustre/iss02/ukbiobank/software"
cd ${wd}

 
 
for proc in CAT12_Volume_adj_probe.rank_norm
do
for model in LR
do
${bind}/qsubshcom "module load R/4.1.2|;
Rscript "minpval_traitsGlobal.R" ${proc} ${model} " 4 12G minTraits_${proc} 04:00:00 "--partition=bigmem"
done
done



for proc in FS_Thickness_adk_probe.rank_norm
do
for model in LR
do
${bind}/qsubshcom "module load R/4.1.2|;
Rscript "minpval_traits_surface.R" ${proc} ${model} " 4 40G minTraits_${proc} 04:00:00 ""
done
done
```

get minimal pvalue distribution per processing across all traits (1000 pval, for each 10000 val)
```{bash, message=FALSE, warning=FALSE,eval=F}
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol/Random_traits/Results"
bind="/network/lustre/iss02/ukbiobank/software"
cd ${wd}

 
 
for proc in FS_Thickness_adj_probe.Rank
do
for model in LR
do
${bind}/qsubshcom "module load R/4.1.2|;
Rscript "Distrib_minpval.R" ${proc} ${model} " 4 12G distribPval_${proc} 04:00:00 ""
done
done


```

#Results 
```{R, message=FALSE, warning=FALSE,eval=F}
library(ggplot2)
library(dplyr)
library(hrbrthemes)


df_anat_lr<-read.table('Results/tab_fin_pval_Anat_LR_1000')
df_anat_lr$V2<-df_anat_lr$V2/10

df_anat_rank_norm_lr<-read.table('Results/tab_fin_pval_Anat_rank_norm_LR_1000')
df_anat_rank_norm_lr$V2<-df_anat_rank_norm_lr$V2/10

df_vbm_lr<-read.table('Results/tab_fin_pval_Vbm_LR_1000')
df_vbm_lr$V2<-df_vbm_lr$V2/10

df_vbm_rank_norm_lr<-read.table('Results/tab_fin_pval_Vbm_rank_norm_LR_1000')
df_vbm_rank_norm_lr$V2<-df_vbm_rank_norm_lr$V2/10

df_cat12_vol_lr<-read.table('Results/tab_fin_pval_CAT12_Volume_LR_1000')
df_cat12_vol_lr$V2<-df_cat12_vol_lr$V2/10

df_cat12_vol_rank_norm_lr<-read.table('Results/tab_fin_pval_CAT12_Volume_rank_norm_LR_1000')
df_cat12_vol_rank_norm_lr$V2<-df_cat12_vol_rank_norm_lr$V2/10


df_cat12_surf_lr<-read.table('Results/tab_fin_pval_CAT12_Surface_LR_1000')
df_cat12_surf_lr$V2<-df_cat12_surf_lr$V2/10

df_cat12_surf_rank_norm_lr<-read.table('Results/tab_fin_pval_CAT12_Surface_rank_norm_LR_1000')
df_cat12_surf_rank_norm_lr$V2<-df_cat12_surf_rank_norm_lr$V2/10

df_fs_thick_lr<-read.table('Results/tab_fin_pval_FS_Thickness_LR_1000')
df_fs_thick_lr$V2<-df_fs_thick_lr$V2/10

df_fs_thick_rank_norm_lr<-read.table('Results/tab_fin_pval_FS_Thickness_rank_norm_LR_1000')
df_fs_thick_rank_norm_lr$V2<-df_fs_thick_rank_norm_lr$V2/10

df_fs_all_lr<-read.table('Results/tab_fin_pval_FS_All_moda_LR_1000')
df_fs_all_lr$V2<-df_fs_all_lr$V2/10

#df_fs_all_rank_norm_lr<-read.table('Results/tab_fin_pval_FS_All_Moda_rank_norm_LR_500')
#df_fs_all_rank_norm_lr$V2<-df_fs_all_rank_norm_lr$V2/5



colnames(df_anat_lr)<-c('alpha','FSLAnat_LR','N')
colnames(df_anat_rank_norm_lr)<-c('alpha','FSLAnat_rank_norm_LR','N')

colnames(df_vbm_lr)<-c('alpha','FSLVbm_LR','N')
colnames(df_vbm_rank_norm_lr)<-c('alpha','FSLVbm_rank_norm_LR','N')

colnames(df_cat12_vol_lr)<-c('alpha','CAT12_Volume_LR','N')
colnames(df_cat12_vol_rank_norm_lr)<-c('alpha','CAT12_Volume_LR','N')

colnames(df_cat12_surf_lr)<-c('alpha','CAT12_Surface_LR','N')
colnames(df_cat12_surf_rank_norm_lr)<-c('alpha','CAT12_Surface_rank_norm_LR','N')

colnames(df_fs_thick_lr)<-c('alpha','FS_Thick_LR','N')
colnames(df_fs_thick_rank_norm_lr)<-c('alpha','FS_Thick_rank_norm_LR','N')

colnames(df_fs_all_lr)<-c('alpha','FS_All_LR','N')


df_anat_lr$Proc<-'FSLANAT'
df_anat_rank_norm_lr$Proc<-'FSLANAT'
df_anat_lr$Method<-'Classic'
df_anat_rank_norm_lr$Method<-'Rank Norm'

df_vbm_lr$Proc<-'FSLVBM'
df_vbm_rank_norm_lr$Proc<-'FSLVBM'
df_vbm_lr$Method<-'Classic'
df_vbm_rank_norm_lr$Method<-'Rank Norm'

df_cat12_vol_lr$Proc<-'CAT12 Volume'
df_cat12_vol_rank_norm_lr$Proc<-'CAT12 Volume'
df_cat12_vol_lr$Method<-'Classic'
df_cat12_vol_rank_norm_lr$Method<-'Rank Norm'


df_cat12_surf_lr$Proc<-'CAT12 Surface'
df_cat12_surf_rank_norm_lr$Proc<-'CAT12 Surface'
df_cat12_surf_lr$Method<-'Classic'
df_cat12_surf_rank_norm_lr$Method<-'Rank Norm'


df_fs_thick_lr$Proc<-'FreeSurfer Cortical Thickness'
df_fs_thick_rank_norm_lr$Proc<-'FreeSurfer Cortical Thickness'
df_fs_thick_lr$Method<-'Classic'
df_fs_thick_rank_norm_lr$Method<-'Rank Norm'

df_fs_all_lr$Proc<-'FreeSurfer All Modalities'
df_fs_all_lr$Method<-'Classic'


colnames(df_anat_lr)<-c('alpha','Nb_FP','N','Proc','Method')
colnames(df_anat_rank_norm_lr)<-c('alpha','Nb_FP','N','Proc','Method')

colnames(df_vbm_lr)<-c('alpha','Nb_FP','N','Proc','Method')
colnames(df_vbm_rank_norm_lr)<-c('alpha','Nb_FP','N','Proc','Method')

colnames(df_cat12_vol_lr)<-c('alpha','Nb_FP','N','Proc','Method')
colnames(df_cat12_vol_rank_norm_lr)<-c('alpha','Nb_FP','N','Proc','Method')

colnames(df_cat12_surf_lr)<-c('alpha','Nb_FP','N','Proc','Method')
colnames(df_cat12_surf_rank_norm_lr)<-c('alpha','Nb_FP','N','Proc','Method')

colnames(df_fs_thick_lr)<-c('alpha','Nb_FP','N','Proc','Method')
colnames(df_fs_thick_rank_norm_lr)<-c('alpha','Nb_FP','N','Proc','Method')

colnames(df_fs_all_lr)<-c('alpha','Nb_FP','N','Proc','Method')


df_tot<-rbind(df_anat_lr,df_anat_rank_norm_lr,df_vbm_lr,df_vbm_rank_norm_lr,df_cat12_vol_lr,df_cat12_vol_rank_norm_lr,df_cat12_surf_lr,df_cat12_surf_rank_norm_lr,df_fs_thick_lr,df_fs_thick_rank_norm_lr,df_fs_all_lr)

#df_tot<-rbind(df_anat_lr,df_vbm_lr,df_cat12_vol_lr,df_cat12_surf_lr,df_fs_thick_lr,df_fs_all_lr)
df_tot$log_alpha<--log10(df_tot$alpha)

df_tot$Nb_FP<-as.numeric(df_tot$Nb_FP)/100
df_tot$Proc<-as.factor(df_tot$Proc)
df_tot$N<-df_tot$N/1000

#nvox_anat=184637
#nvox_vbm=181544
#nvox_cat12=192483
#nvox_cat12_surf=299881
#nvox_fs=654002

df_bonferroni<-df_tot[(df_tot$Proc=="FSLANAT" & df_tot$N==184.637)  | (df_tot$Proc=="FSLVBM" & df_tot$N==181.544) | (df_tot$Proc=="CAT12 Volume" & df_tot$N==192.483) | (df_tot$Proc=="CAT12 Surface" & df_tot$N==299.881) | (df_tot$Proc=="FreeSurfer Cortical Thickness" & df_tot$N==299.881) | (df_tot$Proc=="FreeSurfer All Modalities" & df_tot$N==654.002),]

df_tot$se_high<-df_tot$Nb_FP+0.02*df_tot$Nb_FP #pour 500 traits: sqrt(0.05*0.95/500)*1.96
df_tot$se_low<-df_tot$Nb_FP-0.02*df_tot$Nb_FP

#df_tot$Processing<-factor(df_tot$Proc,levels=c('FSLVBM','FSLVBM_rank_norm','FSLANAT','FSLANAT_rank_norm','CAT12 Volume','CAT12 Volume_rank_norm','CAT12 Surface','CAT12 Surface_rank_norm','FreeSurfer Cortical Thickness','FreeSurfer All Modalities'))

df_tot$Processing<-factor(df_tot$Proc,levels=c('FSLVBM','FSLANAT','CAT12 Volume','CAT12 Surface','FreeSurfer Cortical Thickness','FreeSurfer All Modalities'))

df_tot$Method<-as.factor(df_tot$Method)

#plot: Nb FP en fonction de log alpha pour LR
library(ggplot2)


ggplot(data=df_tot,aes(x=log_alpha,y=Nb_FP,colour=Processing)) +scale_color_manual(values=c('#9CCC65','yellow','#26A69A','#EC407A','#5C6BC0','#512DA8')) +geom_line(linewidth=0.7)+scale_linetype_manual(values=c('solid','dashed'))+ geom_hline(yintercept = 0.05,linewidth=1.5,color='black')+ylab('FWER')+xlab('-log10(alpha)')+geom_point(data=df_bonferroni,aes(x=log_alpha,y=Nb_FP),colour='black',size=3) 


df_tot2<-df_tot[df_tot$Method=='Classic',]
df_bonferroni<-df_bonferroni[df_bonferroni$Method=='Classic',]
ggplot(data=df_tot2,aes(x=log_alpha,y=Nb_FP,colour=Processing)) +scale_color_manual(values=c('#9CCC65','yellow','#26A69A','#EC407A','#5C6BC0','#512DA8')) +geom_line(linewidth=1.4) +geom_point(alpha=0.5,size=2)+ geom_hline(yintercept = 0.05,linewidth=1.5,linetype="dotted",color='black')+ylab('FWER')+xlab('-log10(alpha)')+geom_point(data=df_bonferroni,aes(x=log_alpha,y=Nb_FP),colour='black',size=5) + geom_ribbon(aes(ymin=se_high, ymax=se_low,fill=Processing),alpha = 0.3,linetype=0)+ theme(axis.text.x=element_text(size=30,face="bold"),axis.text.y=element_text(size=30,face="bold"),legend.text =element_text(size=30,face="bold"),axis.title.x = element_text(size=30,face="bold"),axis.title.y = element_text(size=30,face="bold"),legend.title = element_text(size=30,face="bold"))+theme(legend.position=c(0.7,0.8))+scale_fill_manual(values=c('#9CCC65','yellow','#26A69A','#EC407A','#5C6BC0','#512DA8'))+theme(panel.background = element_blank(),panel.grid = element_line(colour="grey"))
```

do it again but --rank normalized
```{bash,eval=F}

wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"
bind="/network/lustre/iss02/ukbiobank/software"
cat12="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesCAT12"
fsl="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFSL"
fs="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFS"

cd ${fsl}

for proc in Anat 
do
for div in main 
do 
${bind}/qsubshcom "${bind}/osca --befile UKB_QC_${proc}.FinalBod.${div}.adj_probe --rint-probe --make-bod --out UKB_QC_${proc}.FinalBod.${div}.adj_probe.rank_norm|;
" 8 150G ${proc}_Norm 120:00:00 "--partition=bigmem"
done
done 



cd ${wd}
for proc in Vbm_rank_norm
do
for div in main 
do 
for seq in $(seq 1 1 500)
do
${bind}/qsubshcom "${bind}/osca --befile ${fsl}/UKB_QC_${proc}.FinalBod.${div} --pheno Random_traits/random_pheno_${div}_${seq}  --covar covar_discr_${div} --qcovar covar2_quant_Vbm_body_size_${div} --linear --out Random_traits/Results/${proc}/resu_${proc}_LR_${div}_${seq}|;
" 4 90G ${div}_${proc}_${seq} 02:30:00 "--partition=bigmem"
done 
done 
done 


#NEW TEST : on adj--probe avant de rank normaliser
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"
bind="/network/lustre/iss02/ukbiobank/software"
cat12="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesCAT12"
fsl="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFSL"
fs="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFS"

cd ${wd}

for proc in CAT12_Volume
do
for div in main 
do 
${bind}/qsubshcom "${bind}/osca --befile ${cat12}/UKB_QC_Vol.FinalBod.${div} --adj-probe --covar covar_discr_main -qcovar covar2_quant_surf_body_size_main --make-bod --out ${cat12}/UKB_QC_${proc}.adj_probe|;
${bind}/osca --befile ${cat12}/UKB_QC_${proc}.adj_probe --rint-probe --make-bod --out ${cat12}/UKB_QC_${proc}.adj_probe.rank_norm|;
" 8 150G ${proc}_Ajd_Probe 120:00:00 "--partition=bigmem"
done
done 


```


where are voxels with min pvalues 
```{R,eval=F}
library(data.table)
AnatAtlas<-fread('../MNI/CODE/all_atlases_Anat')
minpvalAnat<-fread('../../Downloads/pvalmin50_Anat')
colnames(minpvalAnat)<-c('V2','V1')
AnatAtlas50<-AnatAtlas[AnatAtlas$V2 %in% c(minpvalAnat$V2),]
AnatAtlas50<-merge(AnatAtlas50,minpvalAnat,by=c('V2'))
AnatAtlas50<-AnatAtlas50[order(AnatAtlas50$V1.y,decreasing = FALSE),]
summary(as.factor(AnatAtlas50$AtlasH0_3region))
```

where are voxels with min pvalues 
```{R,eval=F}
VbmAtlas<-fread('../MNI/CODE/all_atlases_Vbm')
minpvalVbm<-fread('../../Downloads/pvalmin50_Vbm')

VbmAtlas50<-VbmAtlas[VbmAtlas$V2 %in% c(minpvalVbm$V1),]
summary(as.factor(VbmAtlas50$AtlasH0_3region))
```

plots sumR2+optimal threshold and same with rank norm

```{R,eval=F}
proc<-c('Anat','Vbm','CAT12V','CAT12S','FSTh')
indep<-c(113,81,56,14,53)
rank<-c(112,79,65,14,56)
sumR2<-c(48,42,240,1471,137)
df_indep<-data.frame(indep,sumR2)
colnames(df_indep)<-c('percent_test','sumR2')
df_indep$method<-'classic'
df_indep$Proc<-proc
df_rank<-data.frame(rank,sumR2)
colnames(df_rank)<-c('percent_test','sumR2')
df_rank$method<-'rank_norm'
df_rank$Proc<-proc

df_tot<-rbind(df_indep,df_rank)

library(ggplot2)
ggplot(df_tot,aes(x=percent_test,y=sumR2,color=Proc,shape=method))+geom_point(size=4)

model<-lm(indep~sumR2)
summary(model)
```

Analyzing pvalmin_Traits = voxel with minimal pval for each trait
```{R,eval=F}
library(data.table)
vbm<-fread('Results/pvalmin_Traits_FS_All_moda')
colnames(vbm)<-c('vox','pvalminTrait','sumR2','pvalMinGlobal','Skewness','Kurtosis','logpvalMinGlobal','V1','V3','SubCort_Region','Cort_Region','Cer_Region','AtlasH0_3region','AtlasHO','AtlasJulich','Processing','X','Y','Z','Area')
vbm<-as.data.frame(vbm)
vbm$logpvalTrait<--log10(vbm$pvalminTrait)
vbm_global<-vbm[vbm$Area=='Global',]
vbm_global<-vbm_global[order(vbm_global$logpvalTrait,decreasing=T),]
vbm_global$Trait<-seq(1,1000)


vbm_cort<-vbm[vbm$Area=='Cort',]
vbm_cort<-vbm_cort[order(vbm_cort$logpvalTrait,decreasing=T),]
vbm_cort$Trait<-seq(1,1000)


vbm_subcort<-vbm[vbm$Area=='Subcort',]
vbm_subcort<-vbm_subcort[order(vbm_subcort$logpvalTrait,decreasing=T),]
vbm_subcort$Trait<-seq(1,1000)

vbm_cer<-vbm[vbm$Area=='Cer',]
vbm_cer<-vbm_cer[order(vbm_cer$logpvalTrait,decreasing=T),]
vbm_cer$Trait<-seq(1,1000)

vbm_total10<-rbind(vbm_global[1:10,],vbm_subcort[1:10,],vbm_cort[1:10,],vbm_cer[1:10,])
vbm_total<-rbind(vbm_global,vbm_subcort,vbm_cort,vbm_cer)

ggplot(vbm_total,aes(x=Trait,y=logpvalTrait,colour=Area))+geom_point()



vbm_global[50,'pvalminTrait']
vbm_cort[50,'pvalminTrait']
vbm_subcort[50,'pvalminTrait']
vbm_cer[50,'pvalminTrait']
```

pour adj probe rank norm, anat et vbm regarder en details
```{R,eval=F}
anat_adj<-fread('Results/pvalmin_Traits_Anat')
anat_adj<-as.data.frame(anat_adj)
anat_adj<-anat_adj[order(anat_adj$pval,decreasing=FALSE),]

anat_adj_unique<-anat_adj[,c('vox','pval','AtlasJulich')]
anat_adj_unique<-unique(anat_adj_unique)

anat_adj_unique_global<-anat_adj[anat_adj$Area=='Global',c('vox','pval','AtlasJulich','V1')]
anat_adj_unique_global<-unique(anat_adj_unique_global)
#voxels avec les pval les plus petites : 19811 37291 152165 24469 

#on plot les 1000 sur cerveau

library(oro.nifti)

img<-data.frame(vox=seq(1,902629),val=0)
atlas<-fread('../02_Multiple_testing/df_all_watlas_Anat')
img[img$vox %in% atlas$V1,'val']<-1
img[img$vox %in% anat_adj_unique_global$V1,'val']<-10
arr<-array(img$val,dim=c(91,109,91))
writeNIfTI(arr,'pvalMinTraitAnat_adj_rank')

```


```{bash,eval=F}
bod='/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BodFiles_LDSC'
wd="/network/lustre/iss02/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol/Random_traits/Results"
bind="/network/lustre/iss02/ukbiobank/software"


cd $wd

${bind}/qsubshcom "${bind}/osca --befile ${bod}/UKB_QC_FS_Thickness.adj_probe.rank_norm --extract-probe maxKurt_FSTh --make-efile --out maxKurtFSTh_adj_rank.txt" 1 10G Anat_ajdpr_txt 04:00:00 "--partition=normal" 

${bind}/qsubshcom "${bind}/osca --befile ${bod}/UKB_QC_FS_Thickness_1000.FinalBod.main --extract-probe maxKurt_FSTh --make-efile --out maxKurtFSTh_1000.txt" 1 10G Anat_txt 04:00:00 "--partition=normal" 

```


#on plot leur distrib et on regarde à quoi ces memes voxels ressemblent dans Vbm 
```{R,eval=F}
library(data.table)
Anatmin<-fread('../Filezilla_dl/maxKurtFSTh_1000.txt')
Anatmin<-as.data.frame(Anatmin)
colnames(Anatmin)<-c('ID','ID','probe2','probe3','probe4')
Anatmin<-Anatmin[,-c(1,2)]
Anatmin$Method<-'Basic'

AnatAdjmin<-fread('../Filezilla_dl/maxKurtFSTh_adj_rank.txt')
AnatAdjmin<-as.data.frame(AnatAdjmin)
colnames(AnatAdjmin)<-c('ID','ID','probe2','probe3','probe4')
AnatAdjmin<-AnatAdjmin[,-c(1,2)]
AnatAdjmin$Method<-'Adj_probe_rank_norm'

histo<-rbind(Anatmin,AnatAdjmin)

library(ggplot2)
library(hrbrthemes)
ggplot(histo,aes(x=probe2,fill=Method))+geom_histogram(bins = 300,alpha=0.6)+theme_ipsum()


mi<-fread('Results/pval_min_tot_Vbm')
miadj<-fread('Results/pval_min_tot_Vbm_adj_probe.rank_norm')

```


#Run BWAS on phenotype
```{bash,eval=F}

wd="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"
bind="/network/iss/ukbiobank/software"
bods="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesMain"


cd $wd 

for proc in Vbm FS_All_moda
do
for div in main
do 
${bind}/qsubshcom "pheno=\$(sed -n \"\${TASK_ID}{p;q}\" ${wd}/pheno_allList)|;
echo \${pheno}|;
${bind}/osca --befile ${bods}/UKB_QC_${proc}.FinalBod.main  --pheno df_main_\${pheno} --covar covar_discr_${div} --qcovar covar2_quant_fs_body_size_${div} --linear --out Random_traits/Results/${proc}/resu_${proc}_LR_\${pheno}|;
" 4 150G ${proc} 03:00:00 "-array=1-5"
done 
done

for proc in CAT12_Surface_adj_probe.rank_norm CAT12_Volume_adj_probe.rank_norm 
do
${bind}/qsubshcom "pheno=\$(sed -n \"\${TASK_ID}{p;q}\" ${wd}/pheno_allList)|;
${bind}/osca --befile ${cat12}/UKB_QC_${proc}  --pheno df_main_\${pheno} --linear --out Random_traits/Results/${proc}/resu_${proc}_LR_\${pheno}|;
" 4 90G ${proc} 03:30:00 "-array=1-28"
done 




for proc in Anat Vbm CAT12_Volume CAT12_Surface FS_All_moda
do
for div in main
do 
${bind}/qsubshcom " pheno=\$(sed -n \"\${TASK_ID}{p;q}\" ${wd}/pheno_allList)|;
${bind}/osca --befile ${bods}/UKB_QC_${proc}.FinalBod.main  --pheno df_main_\${pheno} --covar covar_discr_${div} --qcovar covar2_quant_${proc}_${div} --linear --out Random_traits/Results/${proc}/resu_${proc}_LR_\${pheno}|;
" 4 120G ${proc}_BMI 03:00:00 "-array=20-21"
done 
done


for proc in Anat Vbm CAT12_Volume CAT12_Surface FS_All_moda
do
for div in main
do 
${bind}/qsubshcom " pheno=\$(sed -n \"\${TASK_ID}{p;q}\" ${wd}/pheno_allList)|;
${bind}/osca --befile ${bods}/UKB_QC_${proc}.FinalBod.main  --pheno df_main_\${pheno} --covar covar_discr_${div} --qcovar covar_quant_${proc}_${div} --linear --out Random_traits/Results/${proc}/resu_${proc}_LR_\${pheno}|;
" 4 120G ${proc}_BMI 03:00:00 "-array=22-29"
done 
done

```

on test asso entre traits et replication dataset pr voir quels vox répliquent

```{bash,eval=F}
wd="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BRM_Vol"
bind="/network/iss/ukbiobank/software"
fsl="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFSL"
cat12="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesCAT12"
fs="/network/iss/ukbiobank/ukbiobank-remaining/elise.delzant/BodFilesFS"

cd $wd 

for proc in FS_All_moda
do
for div in replicate
do 
${bind}/qsubshcom "pheno=\$(sed -n \"\${TASK_ID}{p;q}\" ${wd}/pheno_allList)|;
echo \${pheno}|;
${bind}/osca --befile ${fs}/UKB_QC_${proc}.FinalBod.${div}  --pheno df_${div}_\${pheno} --covar covar_discr_${div} --qcovar covar2_quant_${proc}_body_size_${div} --linear --out Random_traits/Results/${proc}/resu_${proc}_LR_\${pheno}_${div}|;
" 4 120G ${proc} 08:00:00 "-array=1-18"
done 
done


```
